---
title: "Final Project:"
subtitle: "American Crime and Incarceration"
author: "PS I Love Yue: Oliver, Karam, Steven, Emely"
output: pdf_document
---

```{r packages-data, warning=FALSE, message=FALSE}
library(tidyverse)
crime <- read_csv("data/crime_and_incarceration_by_state.csv") 
```


### Beginning of Project

```{r ex-1}
#Seperate data between Federal prisoners and State prisoners
#Federal observations missing a lot of data so maybe best to work with state

federal_crime <- crime %>%
  filter(jurisdiction == "FEDERAL")

state_crime <- crime %>%
  filter(jurisdiction != "FEDERAL")
```


```{r ex-2}
#population and prisoner proportions

state_crime <- state_crime %>%
  mutate(prisoner_per_pop = prisoner_count/state_population)

state_crime %>%
  arrange(desc(prisoner_per_pop)) %>%
  select(jurisdiction, year, prisoner_count, state_population, prisoner_per_pop)

state_crime %>%
  arrange(state_population) %>%
  select(jurisdiction, year, state_population)

```
We can see here that the 25 highest prisoner per population observations 
are either Alaska or Delaware (two low population states, but not the lowest 
two).

Texas and Oklahoma are the only other states in the top 50 highest prisoner per 
population observations.


```{r ex-3}
#We can use this new data set to look at the most recent data

recent_state_crime <- state_crime %>%
  filter(year == 2016)

recent_state_crime %>%
  arrange(desc(prisoner_per_pop)) %>%
  select(jurisdiction, prisoner_per_pop)

recent_state_crime %>%
  arrange(desc(prisoner_count)) %>%
  select(jurisdiction, prisoner_count)


```
We can see here that in 2016 Oklahoma had the highest proportion of population 
imprisoned, while Texas had the highest raw number.

We also observe that the only states in the top 10 `prisoners_per_pop` and the 
top 10 `prisoner_count` are Texas, Florida, Arizona.




```{r ex-4}

ggplot(data = state_crime, 
       mapping = aes(x = year,
                     y = prisoner_per_pop,
                     color = jurisdiction)) +
  geom_point() + 
  geom_smooth()
```


```{r percent-pop}
crime<- crime %>%
  group_by(year)%>%
  mutate(percent_pop= sum(prisoner_count)/sum(state_population))
  
 crime %>% 
  ggplot(crime, mapping= aes(x=year, y=percent_pop))+
  geom_line()+
  labs(title = "title",
  x = "Year", y = "Percent Population Incarcerated") +
  theme_linedraw()
```


```{r Prisoners%-by-state-2001}
crime<- crime %>%
  filter(year == 2001)%>%
  group_by(jurisdiction)%>%
  mutate(prisoner_2001= prisoner_count/state_population)
  
 crime %>% 
  ggplot(crime, mapping= aes(x= jurisdiction, y=prisoner_2001))+
  geom_col()+
  labs(title = "title",
  x = "Year", y = "prisoners 2001") +
  theme_linedraw()+ 
   theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

"violent_crime_total" is the sum of murder_manslaughter, rape_legacy, robbery, & agg_assault by year and state
"property_crime_total" is the sum of burglary, larceny, & vehicle_theft


```{r total-prisoners}
crime<- crime %>%
  group_by(year)%>%
  mutate(total_prisoner= sum(prisoner_count))
  
 crime %>% 
  ggplot(crime, mapping= aes(x=year, y=total_prisoner))+
  geom_line()+
  labs(title = "title",
  x = "Year", y = "Total US Prisoners") +
  theme_linedraw()
```

```{r total-violent-crime}
crime<- crime %>%
  filter(!is.na(violent_crime_total))%>%
  group_by(year)%>%
  mutate(total_crime= sum(violent_crime_total))
  
 crime %>% 
  ggplot(crime, mapping= aes(x=year, y=total_crime))+
  geom_line()+
  labs(title = "title",
  x = "Year", y = "Total Violent Crime in the US") +
  theme_linedraw()
```

```{r total-violent-crime-by-state}
crime<- crime %>%
  filter(!is.na(violent_crime_total))%>%
  group_by(jurisdiction,year)%>%
  mutate(total_crime= sum(violent_crime_total))
  
 crime %>% 
  ggplot(crime, mapping= aes(x=year, y=total_crime))+
  geom_line()+
   facet_wrap(~jurisdiction)+
  labs(title = "title",
  x = "Year", y = "Total Violent Crime in the US") +
  theme_linedraw()
```

