---
title: "American Crime and Incarceration"
subtitle: "STA 199 Final Project"
author: "Oliver Greenwald, Karam Oubari, Steven Powell, Emely Gutierrez"
output: pdf_document
---

## Synopsis
This research project is aimed to further understand and analyze American prison
incarceration and crime in the 50 states between 2001 and 2016. Analysis was 
done by focusing specifically on state prisoners (as opposed to federal) and the
types of crimes that prisoners had been convicted of. The data set used 
observes 7 types of crimes that can be classified into violent crimes and 
property crimes.


## Introduction and Data

### (The introduction should introduce your general research question and your data (where it came from, how it was collected, what are the cases, what are the variables, etc.).)


The United States currently has the highest prison population in the world at
around 2,094,000. This situation can be traced back to the 1970s, when
politicians like Richard Nixon, Ronald Reagan, and more began the War on Drugs
and their “tough on crime” agendas. Since then, rates of imprisonment have
continued to increase under the guise of “public safety.” However, the reality
is that the US’ crime rates are comparable to countries with similar economies,
yet the rate of incarceration is significantly higher. The US  has only 5% of
the world’s population, and yet it houses 25% of the world’s incarcerated
persons.

In this project we aim to understand the extent of the issue of mass 
incarceration, what sub-statistics can most explain the issue, how the 
numbers connect to public policy, and the future of the issue through basic
predictive modeling.

### Data Description
“The Bureau of Justice Statistics administers the National Prisoners Statistics 
Program (NPS), an annual data collection effort that began in response to a 1926 
congressional mandate. The Uniform Crime Report (UCR) has served as the FBI's 
primary national data collection tool since a 1930 congressional mandate 
directed the Attorney General to ‘acquire, collect, classify, and preserve 
identification, criminal identification, crime, and other records.’ The FBI
collects this information voluntarily submitted by local, state, and federal
law enforcement agencies.”

The user who updated this data set used the raw data from the NPS and wrangled 
it in Python to create the variables and observations. Within the data set there 
are 816 rows and 17 variables. The variables included in the set are US state, 
whether or not the data for that entry includes jails, the year, prisoner count 
on December 31st, whether or not the state in the data entry changed their 
system for reporting crime in comparison to previous years, whether or not crime 
totals are estimated, and numbers of: total state population, violent crimes, 
murders or manslaughter, rapes (using the old definition), rapes (using the new 
definition), robberies, aggravated assaults, property crime, burglaries, 
larceny, and vehicle theft.

\newpage

## Methodology

### (The methodology section should include the variables used to address your research question, as well as any useful visualizations or summary statistics. As well, you should introduce and justify the statistical method(s) that you believe will be useful in answering your research question.)

The following variables, from the original data set, were used to address the 
research questions above:\newline
• `jurisdiction`: a categorical variable for federal jurisdiction and the 50 
states. Federal jurisdiction refers to the legal scope of the government's powers in the United States of America.
\newline
• `year`: a numerical variable for the year of each observation, 2001 through 
2016. \newline 
• `prisoner_count`:a numerical variable for the number of prisoners in each
state for the given year. \newline
• `state_population`: a numerical variable for the population of each state
in a given year. \newline
• `violent_crime_total`: a numerical variable for the total number of violent 
crimes (equals the sum of `murder_manslaughter`, `rape_legacy`, `rape_revised`,
`robbery`, and `agg_assault`). A violent crime is defined as a victim is harmed 
by or threatened with violence. 
\newline
• `murder_manslaughter`: a numerical variable for the numbers of murders and
manslaughters. Manslaughter is defined as an unlawful killing that doesn’t 
involve malice aforethought—intent to seriously harm or kill, or extreme, 
reckless disregard for life.
 \newline
• `rape_legacy`: a numerical variable for the number of rapes (using the old
definition). The old definition of rape is: “The carnal knowledge of a female 
forcibly and against her will”
 \newline
• `rape_revised`: a numerical variable for the number of rapes (using the 
new revised definition). The new definition of rape is: “Penetration no matter 
how slight, of the vagina or anus with any body part or object, or oral 
penetration by a sex organ of anotehr person, without the consent of the victim”
\newline
• `robbery`: a numerical variable for the number of robberies. Robbery is 
defined as the taking of something of value from another person using force 
or violence or the threat of force or violence
\newline
• `agg_assault`: a numerical variable for the number of aggravated assaults. 
Aggravated assault is defined as an attempt to cause serious bodily harm to an 
individual with disregard for human life. Factors that raise an assault to the 
aggravated level typically include the use of a weapon, the status of the 
victim, the intent of the perpetrator, and the degree of injury caused.
\newline
• `property_crime_total`: a numerical variable for the total number of property
crimes (equals the sum of `burglary`, `larceny`, and `vehicle_theft`). In a 
property crime, a victim's property is stolen or destroyed, without the use or
threat of force against the victim.
\newline
• `burglary`: a numerical variable for the number of burglaries. Burglary is 
defined as the unlawful entry into almost any structure (not just a home or 
business) with the intent to commit any crime inside (not just theft/larceny).
No physical breaking and entering is required; the offender may simply trespass 
through an open door. Unlike robbery, which involves use of force or fear to 
obtain another person's property, there is usually no victim present during a 
burglary.
\newline
• `larceny`: a numerical variable for the number of larcenies. Larceny is 
defined as what most people think of as common theft - the taking of someone 
else's property without the use of force.
\newline
• `vehicle_theft`: a numerical variable for the number of vehicle thefts. 
Vehicle theft is defined as motor vehicle theft is defined as the theft or 
attempted theft of a motor vehicle. In the UCR Program, a motor vehicle is a 
self-propelled vehicle that runs on land surfaces and not on rails.


### Statistical Methods
To answer the research questions we are exploring we utilized various 
statistical methods using R. More specifically, we used data visualization, 
and linear regression, to both see the current issue and extrapolate and make
predictions outside the scope of the given data.


### Summary Statistics and Visualizations
First, in this project we aim to understand the extent of the issue of mass
incarceration through summary statistics and visualizations. In Figure 1, we 
visualize the total number of prisoners in the US from 2001 to 2016. As we can 
see the graph shows a steady incline from 2001 to 2008 and then a steady decline 
followed by a sharp decline thereafter. In Figure 2, we visualize the amount of 
crime in the US and notice that it decreases quite steadily between 2001 and 
2016. . In Figure 3.1, we examine the change in the proportion of the population
that is imprisoned between 2001 and 2016. This shows us a steady increase 
between 2001 and 2007 before a steady decrease afterwards. Figure 3.2 shows the 
same information in terms of the percent of a states population that is 
incarcerated in a box plot rather than a line graph. The box plot shows that 
though the median percent of the population does not change significantly from 
year to year, certain years have a larger spread and large outliers. These 
outliers are further explained in Figure 4 where we mapped out the proportion of 
each state's population that was imprisoned in 2016 to get a clearer idea of 
which states contribute most the issue of mass incarceration. Most noticeably, 
Delaware, Oklahoma, and Arizona have have much higher imprisonment proportions 
than the rest of the country. These are likely the outliers in the boxplot from 
Figure 3.2. 
 \newpage

```{r packages-data, echo = FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(broom)
library(sf)
```

```{r state_crime, echo = FALSE, warning=FALSE, message=FALSE}
crime <- read_csv("data/crime_and_incarceration_by_state.csv") 

federal_crime <- crime %>%
  filter(jurisdiction == "FEDERAL")

state_crime <- crime %>%
  filter(jurisdiction != "FEDERAL")
```

```{r total-prisoners, echo = FALSE}
crime <- crime %>%
  group_by(year) %>%
  mutate(total_prisoner_per_year = sum(prisoner_count))
 
crime %>%
  ggplot(crime, mapping = aes(x = year, y = total_prisoner_per_year)) +
  geom_line() +
  labs(title = "Total US Prisoners", 
       subtitle = "2001 to 2016",
       x = "Year", 
       y = "Count",
       caption= "Figure 1")+
  theme_linedraw()
```

```{r total-crime, echo = FALSE}
crime <- crime %>%
  filter(!is.na(violent_crime_total)) %>%
  group_by(year) %>%
  mutate(total_violent_crime = sum(violent_crime_total))

crime <- crime %>%
  filter(!is.na(property_crime_total)) %>%
  group_by(year) %>%
  mutate(total_property_crime= sum(property_crime_total))

crime <- crime %>%
  group_by(year) %>%
  mutate(total_crime = total_property_crime + total_violent_crime)
 
 crime %>%
  ggplot(crime, mapping= aes(x=year, y=total_crime)) +
  geom_line() +
  scale_y_continuous(name="Count", labels = scales::comma) +
  labs(title = "Total Crime in the US",
  subtitle= "2001 to 2016",
  x = "Year", y = "Count",
       caption= "Figure 2")+
  theme_linedraw()
```

```{r percent-pop-linegraph, echo = FALSE}
crime <- crime %>%
  group_by(year)%>%
  mutate(percent_pop_us = 100 * sum(prisoner_count)/sum(state_population))
 
 crime %>%
  ggplot(crime, mapping = aes(x = year, 
                              y = percent_pop_us))+
  geom_line() +
  labs(title = " Percent US Population Incarcerated",
       subtitle = "2001 to 2016",
       x = "Year", 
       y = "Percent",
       caption= "Figure 3.1")+
  theme_linedraw()
```

```{r percent-pop, echo = FALSE}
crime<- crime %>%
  group_by(jurisdiction) %>%
  mutate(percent_state_pop= 100*(prisoner_count/state_population))

 crime %>% 
  ggplot(crime, mapping= aes(x=as.factor(year), y= percent_state_pop))+
  geom_boxplot()+
  labs(title = "Percent State Population Incarcerated",
       subtitle = "2001 to 2016",
  x = "Year", y= "Percent",
       caption= "Figure 3.2")+
  theme_linedraw()
```

```{r prisoner-prop-map, fig.height= 6, fig.width= 11, echo = FALSE}
prop_pris_2016 <- state_crime %>%
  filter(year == 2016) %>%
  mutate(prisoner_per_pop = prisoner_count/state_population) %>%
  mutate(region = tolower(jurisdiction)) %>%
  select(region, prisoner_per_pop)

states_map <- map_data("state")
prop_map_2016 <- left_join(states_map, 
                           prop_pris_2016, 
                           by = "region" )

ggplot(prop_map_2016, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = prisoner_per_pop)) +
  scale_fill_viridis_c(option = "B", direction = -1) + 
  labs(title = "2016 Prisoner Proportions",
       subtitle = "by state",
       fill = "Number of prisoners per person",
       caption= "Figure 4")+
  theme_linedraw()
```

```{r crime-by-region, echo=FALSE, warning = FALSE, message = FALSE}

crime <- crime %>%
  mutate(region = case_when(
  jurisdiction == "MAINE" ~ "Northeast",
  jurisdiction == "NEW HAMPSHIRE" ~ "Northeast",
  jurisdiction == "VERMONT" ~ "Northeast",
  jurisdiction == "NEW YORK" ~ "Northeast",
  jurisdiction == "CONNECTICUT" ~ "Northeast",
  jurisdiction == "NEW JERSEY" ~ "Northeast",
  jurisdiction == "MASSACHUSETTS" ~ "Northeast",
  jurisdiction == "RHODE ISLAND" ~ "Northeast",
  jurisdiction == "PENNSYLVANIA" ~ "Northeast",
  jurisdiction == "NORTH DAKOTA" ~ "North Central",
  jurisdiction == "SOUTH DAKOTA" ~ "North Central",
  jurisdiction == "NEBRASKA" ~ "North Central",
  jurisdiction == "KANSAS" ~ "North Central",
  jurisdiction == "MISSOURI" ~ "North Central",
  jurisdiction == "IOWA" ~ "North Central",
  jurisdiction == "MINNESOTA" ~ "North Central",
  jurisdiction == "WISCONSIN" ~ "North Central",
  jurisdiction == "ILLINOIS" ~ "North Central",
  jurisdiction == "INDIANA" ~ "North Central",
  jurisdiction == "MICHIGAN" ~ "North Central",
  jurisdiction == "OHIO" ~ "North Central",
  jurisdiction == "TEXAS" ~ "South",
  jurisdiction == "OKLAHOMA" ~ "South",
  jurisdiction == "LOUISIANA" ~ "South",
  jurisdiction == "ARKANSAS" ~ "South",
  jurisdiction == "MISSISSIPPI" ~ "South",
  jurisdiction == "ALABAMA" ~ "South",
  jurisdiction == "GEORGIA" ~ "South",
  jurisdiction == "FLORIDA" ~ "South",
  jurisdiction == "TENNESSEE" ~ "South",
  jurisdiction == "SOUTH CAROLINA" ~ "South",
  jurisdiction == "NORTH CAROLINA" ~ "South",
  jurisdiction == "KENTUCKY" ~ "South",
  jurisdiction == "VIRGINIA" ~ "South",
  jurisdiction == "WEST VIRGINIA" ~ "South",
  jurisdiction == "MARYLAND" ~ "South",
  jurisdiction == "DELAWARE" ~ "South",
  jurisdiction == "ALASKA" ~ "West",
  jurisdiction == "HAWAII" ~ "West",
  jurisdiction == "NEW MEXICO" ~ "West",
  jurisdiction == "COLORADO" ~ "West",
  jurisdiction == "WYOMING" ~ "West",
  jurisdiction == "MONTANA" ~ "West",
  jurisdiction == "IDAHO" ~ "West",
  jurisdiction == "UTAH" ~ "West",
  jurisdiction == "ARIZONA" ~ "West",
  jurisdiction == "CALIFORNIA" ~ "West",
  jurisdiction == "NEVADA" ~ "West",
  jurisdiction == "OREGON" ~ "West",
  jurisdiction == "WASHINGTON" ~ "West"))

year_crime<- crime %>%
  group_by(year, region) %>%
  summarise(total_burglary = sum(burglary, na.rm = T),
            total_vehicle_theft = sum(vehicle_theft, na.rm = T),
            total_property_crime = sum(property_crime_total, na.rm = T),
            total_larceny = sum(larceny, na.rm = T))

year_crime <- year_crime %>%
  filter(!is.na(region))


ggplot(year_crime, aes(x=year)) +
  geom_line(aes(y = total_burglary, colour = "Total Burglary")) +
  geom_line(aes(y = total_larceny, colour = "Total Larceny")) +
  geom_line(aes(y = total_vehicle_theft, colour = "Total Vehicle Theft")) +
  geom_line(aes(y = total_property_crime, colour = "Total Property Crime")) +
  scale_colour_manual("",
                      breaks = c("Total Burglary", "Total Larceny", "Total Vehicle Theft", "Total Property Crime"),
                      values = c("red3", "steelblue", "springgreen4", "orangered2")) +
  facet_wrap(~region) +
  scale_y_continuous(name="Number of Crimes", labels = scales::comma) +
  labs(title = " Yearly Crime by Region in the United States", 
       caption = "Figure 5")+
  theme_linedraw()
```

```{r ggplot-of-rape-convictions, echo = FALSE}
crime_rape <- crime
crime_rape[is.na(crime_rape)] <- 0

crime_rape <- crime_rape %>%
  group_by(year)%>%
  mutate(total_rape_legacy= sum(rape_legacy))
 
crime_rape <- crime_rape %>%
  group_by(year)%>%
  mutate(total_rape_revised= sum(rape_revised))
 
  ggplot(data = crime_rape,
       mapping = aes(x = year)) +
    
  geom_line(aes(y = total_rape_legacy), color = "red") +
    geom_line(aes(y = total_rape_revised), color = "blue") +
    geom_vline(xintercept = 2012) +
    labs(title="Revising definition of rape in 2012 led to more rape convictions",
     x = "Year",
     y = "Number of Rape Convictions",
     caption = " 
     Red = Total Number of Prisoners convicted of rape
                      by the Legacy Definition in state custody
                      
                Blue = Total Number of Prisoners convicted of rape
                        by the Revised Definition in state custody
                        
                                                          Figure 6")+ 
    theme_linedraw()
```

```{r policy-colorado, echo = FALSE}
col_crime<- crime %>%
 filter(jurisdiction == "COLORADO")%>%
  mutate(percent_col_pop= 100*(prisoner_count/state_population))
  
 col_crime %>% 
  ggplot(col_crime, mapping= aes(x=year, y= percent_col_pop))+
  geom_line()+
  labs(title = "Percent Colorado State Population Incarcerated",
  x = "Year", y= "Percent Population Incarcerated", 
  caption = "Figure 7.1") +
  theme_linedraw()
```

```{r policy-vermont, echo = FALSE}
ver_crime<- crime %>%
 filter(jurisdiction == "VERMONT")%>%
  mutate(percent_ver_pop= 100*(prisoner_count/state_population))
  
 ver_crime %>% 
  ggplot(ver_crime, mapping= aes(x=year, y= percent_ver_pop))+
  geom_line()+
  labs(title = "Percent Vermont State Population Incarcerated",
  x = "Year", y= "Percent Population Incarcerated", caption= "Figure 7.2") +
  theme_linedraw()
```

```{r policy-washington, echo = FALSE}
was_crime<- crime %>%
 filter(jurisdiction == "WASHINGTON")%>%
  mutate(percent_was_pop= 100*(prisoner_count/state_population))
  
 was_crime %>% 
  ggplot(was_crime, mapping= aes(x=year, y= percent_was_pop))+
  geom_line()+
  labs(title = "Percent Washington State Population Incarcerated",
  x = "Year", y= "Percent Population Incarcerated", caption= "Figure 7.3") +
  theme_linedraw()
```
\newpage

## Results

### (Showcase how you arrived at answers to your question using any techniques we have learned in this class (and some beyond, if you’re feeling adventurous). Provide the main results from your analysis. The goal is not to do an exhaustive data analysis (i.e., do not calculate every statistic and procedure you have learned for every variable), but rather let me know that you are proficient at asking meaningful questions and answering them with results of data analysis, that you are proficient in using R, and that you are proficient at interpreting and presenting the results. Focus on methods that help you begin to answer your research questions.)


### Understanding the extent of the issue of mass incarceration:

Political rhetoric that has come to the forefront in recent decades is the idea 
of law and order, being tough on crime, and ensuring punishment to the fullest 
extent of the law. We went into this project knowing that the number of 
incarcerated individuals in the US is the highest of any other developed country 
with a similar economy and governmental system. Knowing this, we would assume 
that if the US has a strong belief that we need to be tough on crime, the high 
rates of incarceration are due to high crime rates, violent or otherwise. With 
this in mind, we wanted to find if the data would support this hypothesis.

Figure 1 and Figure 2: \newline
What is immediately evident is that even when crime of all types was sharply 
increasing, the number of incarcerated individuals was still steadily 
decreasing. Based on this alone, though we cannot definitively draw this 
conclusion, it would appear that rates of incarceration are actually not due to 
how much or how little crime is being committed and is instead due to other 
factors which we will explore later in our analyses. Possible explanations might 
be policy changes, new sentencing guidelines, amount of policing, and a plethora 
of other reasons.

Figure 3.1, Figure 3.2, and Figure 4: \newline
After drawing the previous conclusion, we thought there may be another 
explanation, even if crime was going down it could be that US population was 
increasing, when we graphed it above, it would appear that the percent of the 
US population which is imprisoned shows very strong similarities to the Total 
US Prison Population graph. However, that being said, even if the population is 
increasing, if the crime rate is staying the same, or decreasing as we observed
previously, there is still no valid reason to incarcerate more people.


### Sub-statistics that can better explain the issue:

How does crime differ for different regions of the US?

Figure 5: \newline
To see how crime differs by region, we made a variable called `region` which 
splits up the United States into 4 distinct regions (North Central, Northeast, 
South, and West). We decided to look at only property crime (burglary, larceny, 
and vehicle theft) and so we plotted the total buglary, total larceny, total 
vehicle theft, and overall total property crime from 2000-2016 on each regional 
graph to see what the trends looked like. 

One positive note with these graphs is that each region has a overall general 
decline in the amount of property crime throughout 2000-2016. This is only a 
speculation, but this would suggest that trends of property crime does 
not differ throughout each region but it does for each crime. 

The south overwhelmingly has the total highest property crime total burglary, 
and total larceny.

An outlier seems to be in the northeast graph in 2015 when there was sharp 
decrease with larceny and burglary, and thus total property crime. There must
have been policy that affected the Northeast only during that time period
that lessened crime

Because of the high number of crimes in the south and the downward spime for 
the northwest graph that the north central and west regions are most similar
in numbers of crimes committed over this 16 year span. 

It is also interesting to see that the types of crimes themselves are 
seemingly related. But this we mean that for each graph, from the highest number
to the lowest number, the crimes are larceny, burglary, then vehicle theft. 

If you were looking at a region to live in the united states and buy a house or 
car, it seems that the best place to do that would be in the North East, 
followed closely by the north central and west region. Thus it would be advised 
to not choose the south, although they are on a decline in property crimes 
committed. 


### How the results and the issue connect to non-statistical public policy

How does policy impact incarceration? This will be explored through the lens of 
changing definition of rape as well as through three states which decriminalized 
marijuana within the time frame of the data set. 

Figure 6: \newline
In 2012 the FBI changed their definition of rape from:
"The old definition was 'The carnal knowledge of a female forcibly and against
her will.' Many agencies interpreted this definition as excluding a long list
of sex offenses that are criminal in most jurisdictions, such as offenses
involving oral or anal penetration, penetration with objects, and rapes of
males. The new Summary definition of Rape is: 'Penetration, no matter how 
slight, of the vagina or anus with any body part or object, or oral penetration 
by a sex organ of another person, without the consent of the victim.'"
 
We wanted to see how the change in definition affected the number of
prisoners convicted for rape. Based on the graph, it seems like after the new
definition was implemented, the number of prisoners who are incarcerated for
rape increased. We will perform tests to find the p-value and try to
determine if this is a correlation relationship or a causal relationship.

This one change in policy had obvious impacts on the number of people 
incarcerated for the crime. While it is only a small example, it supports 
theories on incarceration that emphasize the role of laws and policies on the 
number of people incarcerated rather than the actual number of crimes committed. 

This is assuming a causal relationship between the policy change and the number 
of people incarcerated. The creation and enforcement of policy has a greater 
impact on who is incarcerated and how many people are incarcerated for each 
crime than the actual rate of crime. Despite public perception assuming the 
opposite relationship. Many times the policies can be used to adjust a narrative 
to support a political movement or idea at the time. Crime rates can increase 
when policy is adjusted to include more crime as seen with the change in 
defintion for rape. This is not to say that this change contributed 
significantly to mass incarceration or that the change in definition was unjust. 
Rather, this example in our data exemplifies the impact of policies on 
incarceration rather than crime rate.

Figure 7.1, Figure 7.2, and Figure 7.3: \newline
Colorado and Washington were the first US states to decriminalize the use of 
marijuana for recreational purposes in 2012 and Vermont followed suit in 2013. 
We wanted to see if there would be any notable decline in incarceration in these
three states after recreational legalization. It is unfortunate that drug crimes
were not included in the data set as the War on Drugs is responsible for 30-50% 
of the incarceration that we see today. Nonetheless, we thought we could seek to 
find if there was a notable decline in incarceration after 2012 in Colorado and 
Washington and after 2013 in Vermont. Figures 7.2 and 7.3 for Vermont and 
Washington state, respectively,show pretty strong evidence for a sharp decline 
in incarceration after recreational legalization. Figure 7.1 for Colorado is 
less convincing as the incarceration rate leading up to 2012 was already on the 
decline, however, the incarceration most certainly does not increase with 
recreational legalization, which is definitely not enough to call it a causal 
relationship, however, it is pretty compelling evidence. 



### Future of the issue through basic predictive modeling:

To explore the future of prison incarceration in the United States, we chose to
create a linear model that we could use to extrapolate and make broad 
predictions about years past 2016. In addition, we though it would be 
interesting to compare these predicted numbers to the actual statistics from 
2017, 2018, and 2019. This way we could evaluate the strength of our linear 
model and make a general assessment as to whether it is possible to predict
the future of prison incarceration using past data, or whether these statistics
are better determined by other factors such as public policy changes.

```{r future-proportions, echo = FALSE, message = FALSE, warning = FALSE}
prop_pred <- lm(percent_pop_us ~ year, data = crime)
tidy(prop_pred) %>%
  select(term, estimate)
```
pred. % incarcerated = 5.03374618 - 0.00229865(year)

- This linear model implies that in the year 0, on average, the predicted 
percentage incarcerated in the United States was about 5% (a huge and 
unreasonable extrapolation). It also implies that all else held constant with 
each passing year, the percentage incarcerated in the United States is expected
to decrease by about .002%.
\newline

pred. % incarcerated in 2017 = 5.03374618 - 0.00229865(2017)\newline
pred. % incarcerated in 2017 = `r 5.03374618 - 0.00229865*(2017)`\newline
actual % incarcerated in 2017 = `r 100*(1486000/325100000)` \newline


pred. % incarcerated in 2018 = 5.03374618 - 0.00229865(2018)\newline
pred. % incarcerated in 2018 = `r 5.03374618 - 0.00229865*(2018)`\newline
actual % incarcerated in 2018 = `r 100*(1471200/327200000)` \newline


pred. % incarcerated in 2019 = 5.03374618 - 0.00229865(2019)\newline
pred. % incarcerated in 2019 = `r 5.03374618 - 0.00229865*(2019)`\newline
actual % incarcerated in 2019 = `r 100*(1435500/328200000)` \newline

When compared to the actual data it is clear that this data slightly 
underestimates the percent incarcerated in the United States when 
extrapolated. In addition, it is likely inaccurate to extrapolate further than
just a few years, since it is unlikely that incarceration rates will continue
to drop forever. 


\newpage

## Discussion

### (This section is a conclusion and discussion. This will require a summary of what you have learned about your research question along with statistical arguments supporting your conclusions. Also, critique your own methods and provide suggestions for improving your analysis. Issues pertaining to the reliability and validity of your data and appropriateness of the statistical analysis should also be discussed here. A paragraph on what you would do differently if you were able to start over with the project or what you would do next if you were going to continue work on the project should also be included.)

• If we were to continue with this project we would improve our analysis by 
looking for more political and event-based research to aid our graphs. For 
example, in figure 5 there is an obvious dip in every crime for the northeast 
region in ~2015. We spent a good amount of time trying to research this and see 
if there was any policy implemented that would cause a significant dip or an 
event that would cause crime to go down, but we were unsuccessful. Part of the 
issue was that it may have been one state that caused most of the crime and they 
themselves had something happen in 2015 that caused the entire northeast region 
to dip. But because we put states into regions, it would be tedious, for the 
amount of time we have to finish this project, to calculate the crime for every 
state in it’s own graph and then see if one or a few states caused the dip. So 
we believe that if we had more time, we would eventually be able to figure this 
out to add more concrete observations/knowledge to the discussions of our 
graphs.

• We could test more of the literature on mass incarceration that revolves 
around drug incarcerations if we could find a comparable data set to combine 
with ours. The “War on Drugs” is viewed as one of the drivers of mass 
incarceration in America but data on incarcerations for drug charges are not 
included in this data set. There is also a relationship between drug 
incarcerations and violent crimes that would be interesting to investigate further. 

• Additionally, having access to demographic data among the prison populations 
in the various states would allow us to investigate the racial and socioeconomic 
disparities of mass incarceration. This would be important to understand the 
implications of the population data especially if the data we found was in line 
with most literature that states with “one out of every three Black boys born 
today can expect to be sentenced to prison, compared 1 out 6 Latino boys; one 
out of 17 white boys” (NAACP).

• Considering that this is true population data instead of a sample population, 
it was not possible to perform hypothesis testing or bootstrap sampling. It 
could be interesting to find individual statistics of each conviction over a 
period of time to provide more context to our base rates. However, there is a 
benefit to having the true population data because our results will be reliable. 
In addition, our use of base rates mean we don’t really have a clear picture of 
the individuals and specific cases that are impacted by the issue.

• The data set also has a few discrepancies in how the prison populations are 
reported. Some states included jails in their incarcerated population data while 
others only included prisons. This information is provided in the data under the 
“include_jails” column but we did not factor this into our analysis as there is 
no way to separate the jail and prison populations in the data set. 
Additionally, throughout the data set there are four changes in how crime is 
reported in Alabama and Montana in 2011, New York in 2012, and Georgia in 2015. 
The actual changes in the methods of reporting are not included in the data set 
but recognizing that the way crimes were reported would affect the data. Lastly, 
in Kentucky and Illinois, there are multiple years where the crime totals are 
estimates rather than exact data. This could also affect our data and our 
analysis and should be taken into consideration.

• In terms of the linear regression models, it is also important to note that
these models might be better represented by another type of regression model
such as a logistic one. This was not evaluated in the scope of the project. It
is also relevant to note that the extrapolations for predicting future crime
are extremely rough and aren't likely to accurately predict further than a few
years into the future because previous crime rates are likely less indicative of
future crime than a factor such as policy.


## References

C., Chris. “Crime and Incarceration in the United States: 21st Century State 
Crime and Prison Custody Statistics.” Kaggle, 2018, www.kaggle.com/
christophercorrea/prisoners-and-crime-in-united-states?select=crime_and_
incarceration_by_state.csv. 

Chettiar, Inimai. Why Mass Incarceration Really Is the New Jim Crow. 
American Civil Liberties Union, 8 Mar. 2012, www.aclu.org/blog/racial-
justice/why-mass-incarceration-really-new-jim-crow. 

“Frequently Asked Questions about the Change in the UCR Definition of Rape.” 
Federal Bureau of Investigations , 11 Dec. 2014. 

“Highest to Lowest - Prison Population Total.” World Prison Brief, 
Institute for Crime and Justice Policy Research, www.prisonstudies.org/
highest-to-lowest/prison-population-total?field_region_taxonomy_tid=All. 

Kang-Brown, Jacob, et al. People in Prison in 2019. Vera Institute for 
Justice, May 2020, 
www.vera.org/publications/people-in-prison-in-2019. 

“Mass Incarceration.” American Civil Liberties Union, www.aclu.org/issues/
smart-justice/mass-incarceration. 

“Prison Population Counts.” Bureau of Justice Statistics (BJS), Office of 
Justice Programs, 2019, www.bjs.gov/index.cfm?ty=tp. 

US Census Bureau. “US Population By Zip Code for Both 2000 and 2010.” Kaggle, 
2017, www.kaggle.com/census/us-population-by-zip-code. 
