---
title: "American Crime and Incarceration"
subtitle: "STA 199 Final Project"
author: "Oliver Greenwald, Karam Oubari, Steven Powell, Emely Gutierrez"
output: pdf_document
---

## Synoposis
This research project is aimed to further understand and analyze American prison
incarceration and crime in the 50 states between 2001 and 2016. Analysis was 
done by focusing specifically on state prisoners (as opposed to federal) and the
types of crimes that prisoners had been convicted of. The data set used 
observes 7 types of crimes that can be classified into violent crimes and 
property crimes.


## Introduction and Data

### (The introduction should introduce your general research question and your data (where it came from, how it was collected, what are the cases, what are the variables, etc.).)


The United States currently has the highest prison population in the world at
around 2,094,000. This situation can be traced back to the 1970s, when
politicians like Richard Nixon, Ronald Reagan, and more began the War on Drugs
and their “tough on crime” agendas. Since then, rates of imprisonment have
continued to increase under the guise of “public safety.” However, the reality
is that the US’ crime rates are comparable to countries with similar economies,
yet the rate of incarceration is significantly higher. The US  has only 5% of
the world’s population, and yet it houses 25% of the world’s incarcerated
persons.

In this project we aim to understand the extent of the issue of mass 
incarceration, what sub-statistics can most explain the issue, and how the 
numbers connect to public policy.

### Data Description
“The Bureau of Justice Statistics administers the National Prisoners Statistics 
Program (NPS), an annual data collection effort that began in response to a 1926 
congressional mandate. The Uniform Crime Report (UCR) has served as the FBI's 
primary national data collection tool since a 1930 congressional mandate 
directed the Attorney General to ‘acquire, collect, classify, and preserve 
identification, criminal identification, crime, and other records.’ The FBI
collects this information voluntarily submitted by local, state, and federal
law enforcement agencies.”

The user who updated this data set used the raw data from the NPS and wrangled 
it in Python to create the variables and observations. Within the data set there 
are 816 rows and 17 variables. The variables included in the set are US state, 
whether or not the data for that entry includes jails, the year, prisoner count 
on December 31st, whether or not the state in the data entry changed their 
system for reporting crime in comparison to previous years, whether or not crime 
totals are estimated, and numbers of: total state population, violent crimes, 
murders or manslaughter, rapes (using the old definition), rapes (using the new 
definition), robberies, aggravated assaults, property crime, burglaries, 
larceny, and vehicle theft.

\newpage

## Methodology

### (The methodology section should include the variables used to address your research question, as well as any useful visualizations or summary statistics. As well, you should introduce and justify the statistical method(s) that you believe will be useful in answering your research question.)

The following variables, from the original data set, were used to address the 
research questions above:\newline
• `jurisdiction`: a categorical variable for federal jurisdiction and the 50 
states. \newline
• `year`: a numerical variable for the year of each observation, 2001 through 
2016. \newline 
• `prisoner_count`:a numerical variable for the number of prisoners in each
state for the given year. \newline
• `state_population`: a numerical variable for the population of each state
in a given year. \newline
• `violent_crime_total`: a numerical variable for the total number of violent 
crimes (equals the sum of `murder_manslaughter`, `rape_legacy`, `rape_revised`,
`robbery`, and `agg_assault`). \newline
• `murder_manslaughter`: a numerical variable for the numbers of murders and
manslaughters. \newline
• `rape_legacy`: a numerical variable for the number of rapes (using the old
definition). \newline
• `rape_revised`: a numerical variable for the number of rapes (using the 
new revised definition). \newline
• `robbery`: a numerical variable for the number of robberies. \newline
• `agg_assault`: a numerical variable for the number of aggravated assaults. \newline
• `property_crime_total`: a numerical variable for the total number of property
crimes (equals the sum of `burglary`, `larceny`, and `vehicle_theft`). \newline
• `burglary`: a numerical variable for the number of burglaries. \newline
• `larceny`: a numerical variable for the number of larcenies. \newline
• `vehicle_theft`: a numerical variable for the number of vehicle thefts.

### Statistical Methods
(We need to introduce and justify the statistical method(s) that we believe will 
be useful in answering the research question(s).)


### Summary Statistics and Visualizations
First, in this project we aim to understand the extent of the issue of mass
incarceration through summary statistics and visualizations. In Figure 1, we 
visualize the total number of prisoners in the US from 2001 to 2016. As we can 
see the graph shows a steady incline from 2001 to 2008 and then a steady decline 
followed by a sharp decline thereafter. In Figure 2, we visualize the amount of 
crime in the US and notice that it decreases quite steadily between 2001 and 
2016. In Figure 3, we examine the change in the proportion of the population
that is imprisoned between 2001 and 2016. This shows us a steady increase 
between 2001 and 2007 before a steady decrease afterwards. In Figure 4, we 
mapped out the proportion of each state's population that was imprisoned in 2016
to get a clearer idea of which states contribute most the issue of mass 
incarceration. Most noticeably, Delaware, Oklahoma, and Arizona have have much 
higher imprisonment proportions than the rest of the country.

```{r packages-data, echo = FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(broom)
library(sf)
```

```{r state_crime, echo = FALSE, warning=FALSE, message=FALSE}
crime <- read_csv("data/crime_and_incarceration_by_state.csv") 

federal_crime <- crime %>%
  filter(jurisdiction == "FEDERAL")

state_crime <- crime %>%
  filter(jurisdiction != "FEDERAL")
```

```{r total-prisoners, echo = FALSE}
crime <- crime %>%
  group_by(year) %>%
  mutate(total_prisoner_per_year = sum(prisoner_count))
 
 crime %>%
  ggplot(crime, mapping = aes(x = year, 
                             y = total_prisoner_per_year)) +
  geom_line() +
  labs(title = "Figure 1: Total US Prisoners", 
       subtitle = "between 2001 and 2016",
       x = "Year", 
       y = "Count")
```

```{r total-crime, echo = FALSE}
crime <- crime %>%
  filter(!is.na(violent_crime_total)) %>%
  group_by(year) %>%
  mutate(total_violent_crime = sum(violent_crime_total))

crime <- crime %>%
  filter(!is.na(property_crime_total)) %>%
  group_by(year) %>%
  mutate(total_property_crime= sum(property_crime_total))

crime <- crime %>%
  group_by(year) %>%
  mutate(total_crime = total_property_crime + total_violent_crime)
 
 crime %>%
  ggplot(crime, mapping= aes(x=year, y=total_crime)) +
  geom_line() +
  scale_y_continuous(name="Count", labels = scales::comma) +
  labs(title = "Figure 2: Total Crime in the US",
       subtitle = "from 2001 to 2016",
  x = "Year", y = "Count")
```

```{r percent-pop, echo = FALSE}
crime <- crime %>%
  group_by(year)%>%
  mutate(percent_pop = 100 * sum(prisoner_count)/sum(state_population))
 
 crime %>%
  ggplot(crime, mapping = aes(x = year, 
                              y = percent_pop))+
  geom_line() +
  labs(title = " Figure 3: Percent Population Incarcerated",
       subtitle = "between 2001 and 2016",
       x = "Year", 
       y = "Percent")
```

```{r prisoner-prop-map, echo = FALSE}
prop_pris_2016 <- state_crime %>%
  filter(year == 2016) %>%
  mutate(prisoner_per_pop = prisoner_count/state_population) %>%
  mutate(region = tolower(jurisdiction)) %>%
  select(region, prisoner_per_pop)

states_map <- map_data("state")
prop_map_2016 <- left_join(states_map, 
                           prop_pris_2016, 
                           by = "region" )

ggplot(prop_map_2016, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = prisoner_per_pop)) +
  scale_fill_viridis_c(option = "B", direction=-1) + 
  labs(title = "Figure 4: 2016 Prisoner Proportions",
       subtitle = "by state",
       fill = "Number of prisoners per person")
```

```{r crime-by-region, echo=FALSE, warning = FALSE, message = FALSE}

crime <- crime %>%
  mutate(region = case_when(
  jurisdiction == "MAINE" ~ "Northeast",
  jurisdiction == "NEW HAMPSHIRE" ~ "Northeast",
  jurisdiction == "VERMONT" ~ "Northeast",
  jurisdiction == "NEW YORK" ~ "Northeast",
  jurisdiction == "CONNECTICUT" ~ "Northeast",
  jurisdiction == "NEW JERSEY" ~ "Northeast",
  jurisdiction == "MASSACHUSETTS" ~ "Northeast",
  jurisdiction == "RHODE ISLAND" ~ "Northeast",
  jurisdiction == "PENNSYLVANIA" ~ "Northeast",
  jurisdiction == "NORTH DAKOTA" ~ "North Central",
  jurisdiction == "SOUTH DAKOTA" ~ "North Central",
  jurisdiction == "NEBRASKA" ~ "North Central",
  jurisdiction == "KANSAS" ~ "North Central",
  jurisdiction == "MISSOURI" ~ "North Central",
  jurisdiction == "IOWA" ~ "North Central",
  jurisdiction == "MINNESOTA" ~ "North Central",
  jurisdiction == "WISCONSIN" ~ "North Central",
  jurisdiction == "ILLINOIS" ~ "North Central",
  jurisdiction == "INDIANA" ~ "North Central",
  jurisdiction == "MICHIGAN" ~ "North Central",
  jurisdiction == "OHIO" ~ "North Central",
  jurisdiction == "TEXAS" ~ "South",
  jurisdiction == "OKLAHOMA" ~ "South",
  jurisdiction == "LOUISIANA" ~ "South",
  jurisdiction == "ARKANSAS" ~ "South",
  jurisdiction == "MISSISSIPPI" ~ "South",
  jurisdiction == "ALABAMA" ~ "South",
  jurisdiction == "GEORGIA" ~ "South",
  jurisdiction == "FLORIDA" ~ "South",
  jurisdiction == "TENNESSEE" ~ "South",
  jurisdiction == "SOUTH CAROLINA" ~ "South",
  jurisdiction == "NORTH CAROLINA" ~ "South",
  jurisdiction == "KENTUCKY" ~ "South",
  jurisdiction == "VIRGINIA" ~ "South",
  jurisdiction == "WEST VIRGINIA" ~ "South",
  jurisdiction == "MARYLAND" ~ "South",
  jurisdiction == "DELAWARE" ~ "South",
  jurisdiction == "ALASKA" ~ "West",
  jurisdiction == "HAWAII" ~ "West",
  jurisdiction == "NEW MEXICO" ~ "West",
  jurisdiction == "COLORADO" ~ "West",
  jurisdiction == "WYOMING" ~ "West",
  jurisdiction == "MONTANA" ~ "West",
  jurisdiction == "IDAHO" ~ "West",
  jurisdiction == "UTAH" ~ "West",
  jurisdiction == "ARIZONA" ~ "West",
  jurisdiction == "CALIFORNIA" ~ "West",
  jurisdiction == "NEVADA" ~ "West",
  jurisdiction == "OREGON" ~ "West",
  jurisdiction == "WASHINGTON" ~ "West"))

year_crime<- crime %>%
  group_by(year, region) %>%
  summarise(total_burglary = sum(burglary, na.rm = T),
            total_vehicle_theft = sum(vehicle_theft, na.rm = T),
            total_property_crime = sum(property_crime_total, na.rm = T),
            total_larceny = sum(larceny, na.rm = T))

year_crime <- year_crime %>%
  filter(!is.na(region))


ggplot(year_crime, aes(x=year)) +
  geom_line(aes(y = total_burglary, colour = "Total Burglary")) +
  geom_line(aes(y = total_larceny, colour = "Total Larceny")) +
  geom_line(aes(y = total_vehicle_theft, colour = "Total Vehicle Theft")) +
  geom_line(aes(y = total_property_crime, colour = "Total Property Crime")) +
  scale_colour_manual("",
                      breaks = c("Total Burglary", "Total Larceny", "Total Vehicle Theft", "Total Property Crime"),
                      values = c("red3", "steelblue", "springgreen4", "orangered2")) +
  facet_wrap(~region) +
  scale_y_continuous(name="Number of Crimes", labels = scales::comma) +
  labs(title = " Figure 5: Yearly Crime by Region in the United States")
```
```{r ggplot-of-rape-convictions, echo = FALSE}
crime_rape <- crime
crime_rape[is.na(crime_rape)] <- 0

crime_rape <- crime_rape %>%
  group_by(year)%>%
  mutate(total_rape_legacy= sum(rape_legacy))
 
crime_rape <- crime_rape %>%
  group_by(year)%>%
  mutate(total_rape_revised= sum(rape_revised))
 
  ggplot(data = crime_rape,
       mapping = aes(x = year)) +
    
  geom_line(aes(y = total_rape_legacy), color = "red") +
    geom_line(aes(y = total_rape_revised), color = "blue") +
    geom_vline(xintercept = 2012) +
    labs(title="Figure 6: Revising definition of rape in 2012 led to more rape convictions",
     x = "Year",
     y = "Number of Rape Convictions",
     caption = "Red = Total Number of Prisoners convicted of rape
                      by the Legacy Definition in state custody
                      
                Blue = Total Number of Prisoners convicted of rape
                        by the Revised Definition in state custody")
```
\newpage

## Results

### (Showcase how you arrived at answers to your question using any techniques we have learned in this class (and some beyond, if you’re feeling adventurous). Provide the main results from your analysis. The goal is not to do an exhaustive data analysis (i.e., do not calculate every statistic and procedure you have learned for every variable), but rather let me know that you are proficient at asking meaningful questions and answering them with results of data analysis, that you are proficient in using R, and that you are proficient at interpreting and presenting the results. Focus on methods that help you begin to answer your research questions.)



(We can definitely use the linear regression methods we learned in the last
unit here. Let's discuss this when we meet.)

### Are the increasing incarceration numbers due to increasing crime rates? 

Political rhetoric that has come to the forefront in recent decades is the idea 
of law and order, being tough on crime, and ensuring punishment to the fullest 
extent of the law. We went into this project knowing that the number of 
incarcerated individuals in the US is the highest of any other developed country 
with a similar economy and governmental system. Knowing this, we would assume 
that if the US has a strong belief that we need to be tough on crime, the high 
rates of incarceration are due to high crime rates, violent or otherwise. With 
this in mind, we wanted to find if the data would support this hypothesis.

Figure 1 and 2: \newline
What is immediately evident is that even when crime of all types was sharply 
increasing, the number of incarcerated individuals was still steadily 
decreasing. Based on this alone, though we cannot definitively draw this 
conclusion, it would appear that rates of incarceration are actually not due to 
how much or how little crime is being committed and is instead due to other 
factors which we will explore later in our analyses. Possible explanations might 
be policy changes, new sentencing guidelines, amount of policing, and a plethora 
of other reasons.

Figure 3 and 4: \newline
After drawing the previous conclusion, we thought there may be another 
explanation, even if crime was going down it could be that US population was 
increasing, when we graphed it above, it would appear that the percent of the 
US population which is imprisoned shows very strong similarities to the Total 
US Prison Population graph. However, that being said, even if the population is 
increasing, if the crime rate is staying the same, or decreasing as we observed
previously, there is still no valid reason to incarcerate more people.


### How does crime differ for different regions of the US?

Figure 5: \newline
To see how crime differs by region, we made a variable called `region` which 
splits up the United States into 4 distinct regions (North Central, Northeast, 
South, and West). We decided to look at only property crime (burglary, larceny, 
and vehicle theft) and so we plotted the total buglary, total larceny, total 
vehicle theft, and overall total property crime from 2000-2016 on each regional 
graph to see what the trends looked like. 

One positive note with these graphs is that each region has a overall general 
decline in the amount of property crime throughout 2000-2016. This is only a 
speculation, but this would suggest that trends of property crime does 
not differ throughout each region but it does for each crime. 

The south overwhelmingly has the total highest property crime total burglary, 
and total larceny.

An outlier seems to be in the northeast graph in 2015 when there was sharp 
decrease with larceny and burglary, and thus total property crime. There must
have been policy that affected the Northeast only during that time period
that lessened crime

Because of the high number of crimes in the south and the downward spime for 
the northwest graph that the north central and west regions are most similar
in numbers of crimes committed over this 16 year span. 

It is also interesting to see that the types of crimes themselves are 
seemingly related. But this we mean that for each graph, from the highest number
to the lowest number, the crimes are larceny, burglary, then vehicle theft. 

If you were looking at a region to live in the united states and buy a house or 
car, it seems that the best place to do that would be in the North East, 
followed closely by the north central and west region. Thus it would be advised 
to not choose the south, although they are on a decline in property crimes 
committed. 


### How does policy impact incarceration? This will be explored through the lens of hanging definition of rape.

Figure 6: \newline
In 2012 the FBI changed their definition of rape from:
"The old definition was 'The carnal knowledge of a female forcibly and against
her will.' Many agencies interpreted this definition as excluding a long list
of sex offenses that are criminal in most jurisdictions, such as offenses
involving oral or anal penetration, penetration with objects, and rapes of
males. The new Summary definition of Rape is: 'Penetration, no matter how 
slight, of the vagina or anus with any body part or object, or oral penetration 
by a sex organ of another person, without the consent of the victim.'"
 
We wanted to see how the change in definition affected the number of
prisoners convicted for rape. Based on the graph, it seems like after the new
definition was implemented, the number of prisoners who are incarcerated for
rape increased. We will perform tests to find the p-value and try to
determine if this is a correlation relationship or a causal relationship.

This one change in policy had obvious impacts on the number of people incarcerated for the crime. While it is only a small example, it supports theories on incarceration that emphasize the role of laws and policies on the number of people incarcerated rather than the actual number of crimes committed. This is assuming a causal relationship between the policy change and the number of people incarcerated.
The creation and enforcement of policy has a greater impact on who is incarcerated and how many people are incarcerated for each crime than the actual rate of crime. Despite public perception assuming the opposite relationship. Many times the policies can be used to adjust a narrative to support a political movement or idea at the time. Crime rates can increase when policy is adjusted to include more crime as seen with the change in defintion for rape. This is not to say that this change contributed significantly to mass incarceration or that the change in definition was unjust. Rather, this example in our data exemplifies the impact of policies on incarceration rather than crime rate.

\newpage

```{r creating-proportion-variables, eval = FALSE, echo = FALSE}
crime_props <- crime %>%
  filter(jurisdiction != "FEDERAL") %>%
  mutate(violent_crime_total_prop = (violent_crime_total / prisoner_count)) %>%
  mutate(rape_legacy_prop = (rape_legacy / prisoner_count)) %>%
  mutate(murder_manslaughter_prop = (murder_manslaughter / prisoner_count)) %>%
  mutate(rape_revised_prop = (rape_revised / prisoner_count)) %>%
  mutate(robbery_prop = (robbery / prisoner_count)) %>%
  mutate(agg_asssault_prop = (agg_assault / prisoner_count)) %>%
  mutate(property_crime_total_prop = 
           (property_crime_total / prisoner_count)) %>%
  mutate(burglary_prop = (burglary / prisoner_count)) %>%
  mutate(larceny_prop = (larceny / prisoner_count)) %>%
  mutate(vehicle_theft_prop = (vehicle_theft / prisoner_count))
 
crime_props <- crime_props %>%
  group_by(year) %>%
  mutate(total_rape_legacy_prop = 
           (sum(rape_legacy_prop)/sum(prisoner_count))) %>%
  mutate(total_murder_manslaughter_prop = 
           (sum(murder_manslaughter_prop)/sum(prisoner_count))) %>%
  mutate(total_rape_revised_prop = 
           (sum(rape_revised_prop)/sum(prisoner_count))) %>%
  mutate(total_robbery_prop = (sum(robbery_prop)/sum(prisoner_count))) %>%
  mutate(total_agg_assault_prop = 
           (sum(agg_asssault_prop)/sum(prisoner_count))) %>%
  mutate(total_burglary_prop = (sum(burglary_prop)/sum(prisoner_count))) %>%
  mutate(total_larceny_prop = (sum(larceny_prop)/sum(prisoner_count))) %>%
  mutate(total_vehicle_theft_prop = 
           (sum(vehicle_theft_prop)/sum(prisoner_count))) %>%
  mutate(total_violent_crime_total = 
           (sum(violent_crime_total_prop)/sum(prisoner_count))) %>%
  mutate(total_property_crime_total = 
           (sum(property_crime_total_prop)/sum(prisoner_count)))
 
```
 
 
```{r proportion-of-crimes-including-larcency, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
ggplot(data = crime_props, mapping = aes(x = year)) +
  geom_line(aes(y = total_rape_legacy_prop), color = "red", 
            linetype="twodash") +
  geom_line(aes(y = total_murder_manslaughter_prop), color = "blue", 
            linetype="twodash") +
  geom_line(aes(y = total_rape_revised_prop), color = "green", 
            linetype="twodash") +
  geom_line(aes(y = total_robbery_prop), color = "yellow", 
            linetype="twodash") +
  geom_line(aes(y = total_agg_assault_prop), color = "purple", 
            linetype="twodash") +
  geom_line(aes(y = total_burglary_prop), color = "pink") +
  geom_line(aes(y = total_larceny_prop), color = "orange") +
  geom_line(aes(y = total_vehicle_theft_prop), color = "magenta")
  labs(title = "",
     x = "Year",
     y = "Proportion of prisoners convicted of each crime
            (not including federal incarcerated people)",
     caption = "Red = Proportion of Rape by legacy definiton
                Blue = Proportion of murder/manslaughter
                Green = Proportion of Rape by revised definition
                Yellow = Proportion of robbery
                Purple = Proportion of aggravated assault
                Pink = Proportion of burglary
                Orange = Proportion of larceny
                Magenta = Proportion of vehicle theft")
```

<!-- How do trends in different crimes compare to each other? Do they move 
together or do they change individually? -->


```{r fitted-model, eval = FALSE, echo = FALSE}

violent_crime_total_model <- lm(violent_crime_total ~ year, 
                                data = totals_by_year)
tidy(violent_crime_total_model)


robbery_model <- lm(robbery ~ year, data = totals_by_year)
tidy(robbery_model)


agg_assault_model <- lm(agg_assault ~ year, data = totals_by_year)
tidy(agg_assault_model)


property_crime_total_model <- lm(property_crime_total ~ year, 
                                 data = totals_by_year)
tidy(property_crime_total_model)


burglary_model <- lm(burglary ~ year, data = totals_by_year)
tidy(burglary_model)


larceny_model <- lm(larceny ~ year, data = totals_by_year)
tidy(larceny_model)


vehicle_theft_model <- lm(vehicle_theft ~ year, data = totals_by_year)
tidy(vehicle_theft_model)

```

<!-- Here we attempted to answer whether different types of crime change together -->
<!-- over the years or if different crimes fluctuate in different years. We used  -->
<!-- data from the variables `violent_crime_total`, `robbery`, `agg_assault`, -->
<!-- `property_crime_total`, `burglary`, `larceny`, and `vehicle_theft`. -->

<!-- After attempting to compare the shapes of the crime vs. year graphs all together, -->
<!-- we found that the incredibly different scales make this task impossible to do -->
<!-- on a single graph. We then graphed these different crimes on different graphs -->
<!-- with linear models, making it easier to compare shape but still difficult to  -->
<!-- make objective claims about this data. One notable feature we noticed was that -->
<!-- there was a peak for many crimes between 2004 and 2008. We plan to research  -->
<!-- further to find possible causations for these trends. -->


<!-- Last, we created linear models for each crime to see how they each changed -->
<!-- between 2001 and 2016. Notice, all of these 7 crimes have a negative slope -->
<!-- meaning they overall decreased during this time period. The slopes vary in  -->
<!-- magnitude from about -7000 crimes/year to about -180,000 crimes/year. -->


## Discussion

### (This section is a conclusion and discussion. This will require a summary of what you have learned about your research question along with statistical arguments supporting your conclusions. Also, critique your own methods and provide suggestions for improving your analysis. Issues pertaining to the reliability and validity of your data and appropriateness of the statistical analysis should also be discussed here. A paragraph on what you would do differently if you were able to start over with the project or what you would do next if you were going to continue work on the project should also be included.)

Talk about the lack of observational data. These are already big picture 
statistics that don't tell individual stories. Might have been interesting
to see individual statistics of each conviction over a period of time to 
complement these base rates.

Discuss the inclusion of jails in different states, the change in crime 
reporting and the estimation of crime in some states. 
(`includes_jails`, `crime_reporting_change`, `crimes_estimated`)




## References

https://www.prisonstudies.org/highest-to-lowest/prison-population-total?field_region_taxonomy_tid=All

https://www.aclu.org/blog/racial-justice/why-mass-incarceration-really-new-jim-crow

https://www.aclu.org/issues/smart-justice/mass-incarceration

https://www.kaggle.com/census/us-population-by-zip-code

https://www.bjs.gov/index.cfm?ty=tp&tid=131 

https://www.kaggle.com/christophercorrea/prisoners-and-crime-in-united-states?select=crime_and_incarceration_by_state.csv

https://ucr.fbi.gov/recent-program-updates/new-rape-definition-frequently-asked-questions