---
title: "American Crime and Incarceration"
subtitle: "STA 199 Final Project"
author: "Oliver Greenwald, Karam Oubari, Steven Powell, Emely Gutierrez"
output: pdf_document
---

### Synoposis
This research project is aimed to further understand and analyze American prison
incarceration and crime in the 50 states between 2001 and 2016. 


### Introduction
The United States currently has the highest prison population in the world at 
around 2,094,000. This situation can be traced back to the 1970s, when 
politicians like Richard Nixon, Ronald Reagan, and more began the War on Drugs 
and began promoting “tough on crime” agendas. All this is despite scholars of 
the time believing prisons might have been abolished entirely, as crime and 
imprisonment rates were at an all time low. Mass incarceration has resulted in 
many racial disparities, disproportionately affecting Black men in particular. 
These racial disparities are functionally seen by many as modern-day Jim Crow 
laws which were outlawed in the 1960s.

Rates of imprisonment have continued to increase under the guise of “public 
safety.” However, the reality is that the US’ crime rates are comparable to 
countries with similar economies, yet the rate of incarceration is significantly 
higher. The US has only 5% of the world’s population, and yet it houses 25% of 
the world’s incarcerated persons.

With this context and data set, we are hoping to answer a variety of questions such as the following :

1) Is there a relationship between proportion of population incarcerated and party affiliation in American states?
2) How does policy impact incarceration? This will be explored through the lens of changing definition of rape
3) How do trends in different crimes compare to each other? Do they move together or do they change individually?
4) How does crime differ for different regions of the US?



### Data Description
“The Bureau of Justice Statistics administers the National Prisoners Statistics 
Program (NPS), an annual data collection effort that began in response to a 1926 
congressional mandate. The Uniform Crime Report (UCR) has served as the FBI's 
primary national data collection tool since a 1930 congressional mandate 
directed the Attorney General to ‘acquire, collect, classify, and preserve 
identification, criminal identification, crime, and other records.’ The FBI
collects this information voluntarily submitted by local, state, and federal
law enforcement agencies.” The user who updated this dataset used the raw data 
from the NPS and wrangled it in python to create the variables and observations. 
Within the data set there are 816 rows and 17 variables. The variables included 
in the set are US state, whether or not the data for that entry includes jails, 
the year, prisoner count on December 31st, whether or not the state in the data 
entry changed their system for reporting crime in comparison to previous years, 
whether or not crime totals are estimated, and numbers of: total state 
population, violent crimes, murders or manslaughters, rapes (using the old 
definition), rapes (using the new definition), robberies, aggravated assaults, 
property crime, burglaries, larceny, and vehicle theft. 

### Methodology

The following variables, from the original dataset, were used to address the 
research questions above:\newline
• `jurisdiction`: a categorical variable for federal jurisdiction and the 50 
states. \newline
• `year`: a numerical variable for the year of each observation, 2001 through 
2016. \newline 
• `prisoner_count`:a numerical variable for the number of prisoners. \newline
• `state_population`: a numerical variable for the population of each state
in a given year. \newline
• `violent_crime_total`: a numerical variable for the number of violent cr\newline
• `murder_manslaughter`:\newline
• `rape_legacy`:\newline
• `rape_revised`:\newline
• `robbery`:\newline
• `agg_assault`:\newline
• `property_crime_total`:\newline
• `burglary`:\newline
• `larceny`:\newline
• `vehicle_theft`:\newline
• `total_prisoner`:\newline
• `total_violent_crime`:\newline
• `total_crime`:\newline

```{r packages-data, echo = FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(broom)
library(sf)
```

```{r state_crime, echo = FALSE, warning=FALSE, message=FALSE}
crime <- read_csv("data/crime_and_incarceration_by_state.csv") 

federal_crime <- crime %>%
  filter(jurisdiction == "FEDERAL")

state_crime <- crime %>%
  filter(jurisdiction != "FEDERAL")
```

```{r prisoner-prop-map, echo = FALSE}
prop_pris_2001 <- state_crime %>%
  filter(year == 2001) %>%
  mutate(prisoner_per_pop = prisoner_count/state_population) %>%
  mutate(region = tolower(jurisdiction)) %>%
  select(region, prisoner_per_pop)

states_map <- map_data("state")
prop_map_2001 <- left_join(states_map, 
                           prop_pris_2001, 
                           by = "region" )

ggplot(prop_map_2001, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = prisoner_per_pop)) +
  scale_fill_viridis_c(option = "B", direction=-1) + 
  labs(title = "2001 Prisoner Proportions",
       fill = "Number of prisoners per person")

prop_pris_2016 <- state_crime %>%
  filter(year == 2016) %>%
  mutate(prisoner_per_pop = prisoner_count/state_population) %>%
  mutate(region = tolower(jurisdiction)) %>%
  select(region, prisoner_per_pop)

states_map <- map_data("state")
prop_map_2016 <- left_join(states_map, 
                           prop_pris_2016, 
                           by = "region" )

ggplot(prop_map_2016, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = prisoner_per_pop)) +
  scale_fill_viridis_c(option = "B", direction=-1) + 
  labs(title = "2016 Prisoner Proportions",
       fill = "Number of prisoners per person")
```



### Research Question #1 
### Is there a relationship between proportion of population incarcerated and party affiliation in American states?

Political rhetoric that has come to the forefront in recent decades is the idea 
of law and order, being tough on crime, and ensuring punishment to the fullest 
extent of the law. We went into this project knowing that the number of 
incarcerated individuals in the US is the highest of any other developed country 
with a similar economy and governmental system. Knowing this, we would assume 
that if the US has a strong belief that we need to be tough on crime, the high 
rates of incarceration are due to high crime rates, violent or otherwise. With 
this in mind, we wanted to find if the data would support this hypothesis.

```{r total-prisoners, echo = FALSE}
crime <- crime %>%
  group_by(year)%>%
  mutate(total_prisoner= sum(prisoner_count))
 
 crime %>%
  ggplot(crime, mapping= aes(x=year, y=total_prisoner))+
  geom_line()+
  labs(title = "Total US Prisoners", 
       subtitle = "between 2001 and 2016",
       x = "Year", 
       y = "Count") +
  theme_linedraw()
```
Above is the total prisoners in all 50 US states from 2001 to 2016. The graph 
shows a steady incline from 2001 to 2008 and then a steady decline followed by a 
sharp decline thereafter. Drawing attention to the range of the y-axis, we can 
see that despite these sharp declines, the number of incarcerated individuals at 
the lowest point was still above 1.19 million, which is still much higher 
than the rate of incarceration in other countries.

```{r total-violent-crime, echo = FALSE}
crime <- crime %>%
  filter(!is.na(violent_crime_total))%>%
  group_by(year)%>%
  mutate(total_violent_crime= sum(violent_crime_total))
 
 crime %>%
  ggplot(crime, mapping= aes(x=year, y=total_violent_crime))+
  geom_line()+
  labs(title = "Total Violent Crime in the US",
  x = "Year", y = "Count") +
  theme_linedraw()
```

```{r total-property-crime, echo = FALSE}
crime <- crime %>%
  filter(!is.na(property_crime_total))%>%
  group_by(year)%>%
  mutate(total_property_crime= sum(property_crime_total))
 
 crime %>%
  ggplot(crime, mapping= aes(x=year, y=total_property_crime))+
  geom_line()+
  labs(title = "Total Property Crime in the US", 
       subtitle = "Sum of Property Crime and Violent Crime", 
       x = "Year", 
       y = "Count") +
  theme_linedraw()
```


```{r total-crime, echo = FALSE}
crime<- crime %>%
  group_by(year)%>%
  mutate(total_crime= total_property_crime + total_violent_crime)
 
 crime %>%
  ggplot(crime, mapping= aes(x=year, y=total_crime))+
  geom_line()+
  scale_y_continuous(name="Count", labels = scales::comma) +
  labs(title = "Total Crime in the US",
  x = "Year", y = "Count") +
  theme_linedraw()
```

The three graphs above show the change in total violent crime 
(defined as murder/manslaughter, robbery, rape, and aggravated assault), 
total property crime (defined as burglary, larceny, and vehicle theft) and the 
total crime (sum of property and violent crime) committed between 2001 and 2016. 
What is immediately evident is that even when crime of all types was sharply 
decreasing, the number of incarcerated individuals was still steadily 
decreasing. Based on this alone, though we cannot definitively draw this 
conclusion, it would appear that rates of incarceration are actually not due to 
how much or how little crime is being committed and is instead due to other 
factors which we will explore later in our analyses. Possible explanations might 
be policy changes, new sentencing guidelines, amount of policing, and a plethora 
of other reasons.


###Wouldn't this show a correlation between decreasing incarceration and decreasing crime rates?



```{r percent-pop, echo = FALSE}
crime<- crime %>%
  group_by(year)%>%
  mutate(percent_pop= sum(prisoner_count)/sum(state_population))
 
 crime %>%
  ggplot(crime, mapping= aes(x=year, y=percent_pop))+
  geom_line()+
  labs(title = "Percent Population Incarcerated",
  x = "Year", y = "Count") +
  theme_linedraw()
```
After drawing the previous conclusion, we thought there may be another 
explanation, even if crime was going down it could be that US population was 
increasing, when we graphed it above, it would appear that the percent of the 
US population which is imprisoned shows very strong similarities to the Total 
US Prison Population graph. However, that being said, even if the population is 
increasing, if the crime rate is staying the same, or decreasing as we observed
previously, there is still no valid reason to incarcerate more people.

###I feel like this doesn't answer the question above, so change the question or add political stuff?

### Research Question #2 
### How does policy impact incarceration? This will be explored through the lens of changing definition of rape

### Impact of Changing the Definition of Rape
```{r creation-of-rape-data-set, echo = FALSE}
crime_rape <- crime
crime_rape[is.na(crime_rape)] <- 0
```
 
```{r ggplot-of-rape-convictions, echo = FALSE}
crime_rape<- crime_rape %>%
  group_by(year)%>%
  mutate(total_rape_legacy= sum(rape_legacy))
 
crime_rape<- crime_rape %>%
  group_by(year)%>%
  mutate(total_rape_revised= sum(rape_revised))
 
  ggplot(data = crime_rape,
       mapping = aes(x = year)) +
  geom_line(aes(y = total_rape_legacy), color = "red") +
    geom_line(aes(y = total_rape_revised), color = "blue") +
    geom_vline(xintercept = 2012) +
    labs(title = "",
     x = "Year",
     y = "Number of Rape Cases",
     caption = "Red = Total Number of Prisoners convicted of rape
                      by the Legacy Definition in state custody
                Blue = Total Number of Prisoners convicted of rape
                        by the Revised Definition in state custody")
```
In 2012 the FBI changed their definition of rape from:
"The old definition was 'The carnal knowledge of a female forcibly and against
her will.' Many agencies interpreted this definition as excluding a long list
of sex offenses that are criminal in most jurisdictions, such as offenses
involving oral or anal penetration, penetration with objects, and rapes of
males.
 
to:
The new Summary definition of Rape is: 'Penetration, no matter how slight, of
the vagina or anus with any body part or object, or oral penetration by a sex
organ of another person, without the consent of the victim.'"
 
We wanted to see how the change in definition affected the number of
prisoners convicted for rape. Based on the graph, it seems like after the new
definition was implemented, the number of prisoners who are incarcerated for
rape increased. We will perform tests to find the p-value and try to
determine if this is a correlation relationship or a causal relationship.
 
https://ucr.fbi.gov/recent-program-updates/new-rape-definition-
frequently-asked-questions
 
### Proportions of Crime in the United States
 
```{r creating-proportion-variables, echo = FALSE}
crime_props <- crime %>%
  filter(jurisdiction != "FEDERAL") %>%
  mutate(violent_crime_total_prop = (violent_crime_total / prisoner_count)) %>%
  mutate(rape_legacy_prop = (rape_legacy / prisoner_count)) %>%
  mutate(murder_manslaughter_prop = (murder_manslaughter / prisoner_count)) %>%
  mutate(rape_revised_prop = (rape_revised / prisoner_count)) %>%
  mutate(robbery_prop = (robbery / prisoner_count)) %>%
  mutate(agg_asssault_prop = (agg_assault / prisoner_count)) %>%
  mutate(property_crime_total_prop = (property_crime_total / prisoner_count)) %>%
  mutate(burglary_prop = (burglary / prisoner_count)) %>%
  mutate(larceny_prop = (larceny / prisoner_count)) %>%
  mutate(vehicle_theft_prop = (vehicle_theft / prisoner_count))
 
crime_props <- crime_props %>%
  group_by(year) %>%
  mutate(total_rape_legacy_prop = (sum(rape_legacy_prop)/sum(prisoner_count))) %>%
  mutate(total_murder_manslaughter_prop = (sum(murder_manslaughter_prop)/sum(prisoner_count))) %>%
  mutate(total_rape_revised_prop = (sum(rape_revised_prop)/sum(prisoner_count))) %>%
  mutate(total_robbery_prop = (sum(robbery_prop)/sum(prisoner_count))) %>%
  mutate(total_agg_assault_prop = (sum(agg_asssault_prop)/sum(prisoner_count))) %>%
  mutate(total_burglary_prop = (sum(burglary_prop)/sum(prisoner_count))) %>%
  mutate(total_larceny_prop = (sum(larceny_prop)/sum(prisoner_count))) %>%
  mutate(total_vehicle_theft_prop = (sum(vehicle_theft_prop)/sum(prisoner_count))) %>%
  mutate(total_violent_crime_total = (sum(violent_crime_total_prop)/sum(prisoner_count))) %>%
  mutate(total_property_crime_total = (sum(property_crime_total_prop)/sum(prisoner_count)))
 
```
 
```{r proportion-of-crimes-including-larcency, echo = FALSE, warning = FALSE, message = FALSE,}
ggplot(data = crime_props, mapping = aes(x = year)) +
  geom_line(aes(y = total_rape_legacy_prop), color = "red",linetype="twodash") +
  geom_line(aes(y = total_murder_manslaughter_prop), color = "blue",linetype="twodash") +
  geom_line(aes(y = total_rape_revised_prop), color = "green",linetype="twodash") +
  geom_line(aes(y = total_robbery_prop), color = "yellow",linetype="twodash") +
  geom_line(aes(y = total_agg_assault_prop), color = "purple",linetype="twodash") +
  geom_line(aes(y = total_burglary_prop), color = "pink") +
  geom_line(aes(y = total_larceny_prop), color = "orange") +
  geom_line(aes(y = total_vehicle_theft_prop), color = "magenta")
  labs(title = "",
     x = "Year",
     y = "Proportion of prisoners convicted of each crime
            (not including federal incarcerated people)",
     caption = "Red = Proportion of Rape by legacy definiton
                Blue = Proportion of murder/manslaughter
                Green = Proportion of Rape by revised definition
                Yellow = Proportion of robbery
                Purple = Proportion of aggravated assault
                Pink = Proportion of burglary
                Orange = Proportion of larceny
                Magenta = Proportion of vehicle theft")
```

###STEVEN CAN YOU FIX THIS GRAPH TO NOT HAVE COMMENT LOOK WEIRD?

```{r proportion-of-crimes-including-larcency-2, echo = FALSE, warning=FALSE, message=FALSE}
ggplot(data = crime_props, mapping = aes(x = year)) +
  geom_line(aes(y = total_rape_legacy_prop), color = "red", linetype="twodash") +
  geom_line(aes(y = total_murder_manslaughter_prop), color = "blue",linetype="twodash") +
  geom_line(aes(y = total_rape_revised_prop), color = "green",linetype="twodash") +
  geom_line(aes(y = total_robbery_prop), color = "yellow",linetype="twodash") +
  geom_line(aes(y = total_agg_assault_prop), color = "purple",linetype="twodash") +
  geom_line(aes(y = total_burglary_prop), color = "pink") +
  geom_line(aes(y = total_vehicle_theft_prop), color = "magenta") +
  labs(title = "",
     x = "Year",
     y = "Proportion of prisoners convicted of each crime
            (not including federal incarcerated people)",
     caption = "Red = Proportion of Rape by legacy definiton
                Blue = Proportion of murder/manslaughter
                Green = Proportion of Rape by revised definition
                Yellow = Proportion of robbery
                Purple = Proportion of aggravated assault
                Pink = Proportion of burglary
                Magenta = Proportion of vehicle theft")
```
 
```{r plot-violent-nonviolent-crimes, echo = FALSE}
ggplot(data = crime_props, mapping = aes(x = year)) +
  geom_line(aes(y = total_violent_crime_total), color = "red") +
  geom_line(aes(y = total_property_crime_total), color = "blue") +
  labs(title = "",
     x = "Year",
     y = "Proportion of prisoners convicted of each type of crime
            (not including federal incarcerated people)")
```
We wanted to see how the proportions of crimes were divided among the
prior population. Also we wanted to see whether violent or non-violent crimes
were represented more among the prison population. It appears that non violent
crimes make up a larger proportion of the prison population than violent crimes.
Generally, violent and non-violent crimes are decreasing in regards to 
proportion of the prison population. Nonviolent crimes seem to be decreasing at 
a faster rate than violent crimes. Larceny makes up the largest proportion of 
crimes in the prison population. So we took it out for the second visualization
to get a better look at the other crimes. Murder and manslaughter make up the
smallest proportion.

###How does what you did tie into the original rape question you were finding answers to?

### Research Question #3 
### How do trends in different crimes compare to each other? Do they move together or do they change individually?

```{r new-dataset, echo = FALSE, warning=FALSE, message=FALSE}
totals_by_year <- state_crime %>%
  group_by(year) %>%
  summarise(violent_crime_total = sum(violent_crime_total),
            robbery = sum(robbery),
            agg_assault = sum(agg_assault),
            property_crime_total = sum(property_crime_total),
            burglary = sum(burglary),
            larceny = sum(larceny),
            vehicle_theft = sum(vehicle_theft))

# for some reason 2015 didn't get added to my dataset
```


```{r line-graph, echo = FALSE}


ggplot(data = totals_by_year,
       aes(x = year)) +
  geom_line(aes(y = violent_crime_total), color = "red") +
  geom_line(aes(y = robbery), color = "blue") +
  geom_line(aes(y = agg_assault), color = "green") +
  geom_line(aes(y = property_crime_total), color = "yellow") +
  geom_line(aes(y = burglary), color = "pink") +
  geom_line(aes(y = larceny), color = "black") +
  geom_line(aes(y = vehicle_theft), color = "purple")

#not a very effective visualization
  
```


```{r graphs, warning = FALSE, message = FALSE, echo = FALSE}

ggplot(data = totals_by_year, aes(x = year, y = violent_crime_total)) +
  geom_line() +
  geom_smooth(method = lm)

ggplot(data = totals_by_year, aes(x = year, y = robbery)) +
  geom_line() +
  geom_smooth(method = lm)

ggplot(data = totals_by_year, aes(x = year, y = agg_assault)) +
  geom_line() +
  geom_smooth(method = lm)

ggplot(data = totals_by_year, aes(x = year, y = property_crime_total)) +
  geom_line() +
  geom_smooth(method = lm)

ggplot(data = totals_by_year, aes(x = year, y = burglary)) +
  geom_line() +
  geom_smooth(method = lm)

ggplot(data = totals_by_year, aes(x = year, y = larceny)) +
  geom_line() +
  geom_smooth(method = lm)

ggplot(data = totals_by_year, aes(x = year, y = vehicle_theft)) +
  geom_line() +
  geom_smooth(method = lm)

```



```{r fitted-model, echo = FALSE}

violent_crime_total_model <- lm(violent_crime_total ~ year, 
                                data = totals_by_year)
tidy(violent_crime_total_model)


robbery_model <- lm(robbery ~ year, data = totals_by_year)
tidy(robbery_model)


agg_assault_model <- lm(agg_assault ~ year, data = totals_by_year)
tidy(agg_assault_model)


property_crime_total_model <- lm(property_crime_total ~ year, 
                                 data = totals_by_year)
tidy(property_crime_total_model)


burglary_model <- lm(burglary ~ year, data = totals_by_year)
tidy(burglary_model)


larceny_model <- lm(larceny ~ year, data = totals_by_year)
tidy(larceny_model)


vehicle_theft_model <- lm(vehicle_theft ~ year, data = totals_by_year)
tidy(vehicle_theft_model)

```

Here we attempted to answer whether different types of crime change together
over the years or if different crimes fluctuate in different years. We used 
data from the variables `violent_crime_total`, `robbery`, `agg_assault`,
`property_crime_total`, `burglary`, `larceny`, and `vehicle_theft`.

After attempting to compare the shapes of the crime vs. year graphs all together,
we found that the incredibly different scales make this task impossible to do
on a single graph. We then graphed these different crimes on different graphs
with linear models, making it easier to compare shape but still difficult to 
make objective claims about this data. One notable feature we noticed was that
there was a peak for many crimes between 2004 and 2008. We plan to research 
further to find possible causations for these trends.

###Do the research^^?

Last, we created linear models for each crime to see how they each changed
between 2001 and 2016. Notice, all of these 7 crimes have a negative slope
meaning they overall decreased during this time period. The slopes vary in 
magnitude from about -7000 crimes/year to about -180,000 crimes/year.


### Research Question #4
### How does crime differ for different regions of the US?
```{r crime-by-region, echo=FALSE, warning = FALSE, message = FALSE}

crime <- crime %>%
  mutate(region = case_when(
  jurisdiction == "MAINE" ~ "Northeast",
  jurisdiction == "NEW HAMPSHIRE" ~ "Northeast",
  jurisdiction == "VERMONT" ~ "Northeast",
  jurisdiction == "NEW YORK" ~ "Northeast",
  jurisdiction == "CONNECTICUT" ~ "Northeast",
  jurisdiction == "NEW JERSEY" ~ "Northeast",
  jurisdiction == "MASSACHUSETTS" ~ "Northeast",
  jurisdiction == "RHODE ISLAND" ~ "Northeast",
  jurisdiction == "PENNSYLVANIA" ~ "Northeast",
  jurisdiction == "NORTH DAKOTA" ~ "North Central",
  jurisdiction == "SOUTH DAKOTA" ~ "North Central",
  jurisdiction == "NEBRASKA" ~ "North Central",
  jurisdiction == "KANSAS" ~ "North Central",
  jurisdiction == "MISSOURI" ~ "North Central",
  jurisdiction == "IOWA" ~ "North Central",
  jurisdiction == "MINNESOTA" ~ "North Central",
  jurisdiction == "WISCONSIN" ~ "North Central",
  jurisdiction == "ILLINOIS" ~ "North Central",
  jurisdiction == "INDIANA" ~ "North Central",
  jurisdiction == "MICHIGAN" ~ "North Central",
  jurisdiction == "OHIO" ~ "North Central",
  jurisdiction == "TEXAS" ~ "South",
  jurisdiction == "OKLAHOMA" ~ "South",
  jurisdiction == "LOUISIANA" ~ "South",
  jurisdiction == "ARKANSAS" ~ "South",
  jurisdiction == "MISSISSIPPI" ~ "South",
  jurisdiction == "ALABAMA" ~ "South",
  jurisdiction == "GEORGIA" ~ "South",
  jurisdiction == "FLORIDA" ~ "South",
  jurisdiction == "TENNESSEE" ~ "South",
  jurisdiction == "SOUTH CAROLINA" ~ "South",
  jurisdiction == "NORTH CAROLINA" ~ "South",
  jurisdiction == "KENTUCKY" ~ "South",
  jurisdiction == "VIRGINIA" ~ "South",
  jurisdiction == "WEST VIRGINIA" ~ "South",
  jurisdiction == "MARYLAND" ~ "South",
  jurisdiction == "DELAWARE" ~ "South",
  jurisdiction == "ALASKA" ~ "West",
  jurisdiction == "HAWAII" ~ "West",
  jurisdiction == "NEW MEXICO" ~ "West",
  jurisdiction == "COLORADO" ~ "West",
  jurisdiction == "WYOMING" ~ "West",
  jurisdiction == "MONTANA" ~ "West",
  jurisdiction == "IDAHO" ~ "West",
  jurisdiction == "UTAH" ~ "West",
  jurisdiction == "ARIZONA" ~ "West",
  jurisdiction == "CALIFORNIA" ~ "West",
  jurisdiction == "NEVADA" ~ "West",
  jurisdiction == "OREGON" ~ "West",
  jurisdiction == "WASHINGTON" ~ "West"))

year_crime<- crime %>%
  group_by(year, region) %>%
  summarise(total_burglary = sum(burglary, na.rm = T),
            total_vehicle_theft = sum(vehicle_theft, na.rm = T),
            total_property_crime = sum(property_crime_total, na.rm = T),
            total_larceny = sum(larceny, na.rm = T))

year_crime <- year_crime %>%
  filter(!is.na(region))


ggplot(year_crime, aes(x=year)) +
  geom_line(aes(y = total_burglary, colour = "Total Burglary")) +
  geom_line(aes(y = total_larceny, colour = "Total Larceny")) +
  geom_line(aes(y = total_vehicle_theft, colour = "Total Vehicle Theft")) +
  geom_line(aes(y = total_property_crime, colour = "Total Property Crime")) +
  scale_colour_manual("",
                      breaks = c("Total Burglary", "Total Larceny", "Total Vehicle Theft", "Total Property Crime"),
                      values = c("red3", "steelblue", "springgreen4", "orangered2")) +
  facet_wrap(~region) +
  scale_y_continuous(name="Number of Crimes", labels = scales::comma) +
  labs(title = "Yearly Crime by Region in the United States")

view(crime)

```
To see how crime differs by region, we made a variable called `region` which 
splits up the United States into 4 distinct regions (North Central, Northeast, 
South, and West). We decided to look at only property crime (burglary, larceny, 
and vehicle theft) and so we plotted the total buglary, total larceny, total 
vehicle theft, and overall total property crime from 2000-2016 on each regional 
graph to see what the trends looked like. 

One positive note with these graphs is that each region has a overall general 
decline in the amount of property crime throughout 2000-2016. This is only a 
speculation, but this would suggest that trends of property crime does 
not differ throughout each region but it does for each crime. 

The south overwhelmingly has the total highest property crime total burglary, 
and total larceny.

An outlier seems to be in the northeast graph in 2015 when there was sharp 
decrease with larceny and burglary, and thus total property crime. There must
have been policy that affected the Northeast only during that time period
that lessened crime

Because of the high number of crimes in the south and the downward spime for 
the northwest graph that the north central and west regions are most similar
in numbers of crimes committed over this 16 year span. 

It is also interesting to see that the types of crimes themselves are 
seemingly related. But this we mean that for each graph, from the highest number
to the lowest number, the crimes are larceny, burglary, then vehicle theft. 

If you were looking at a region to live in the united states and buy a house or 
car, it seems that the best place to do that would be in the North East, 
followed closely by the north central and west region. Thus it would be advised 
to not choose the south, although they are on a decline in property crimes 
committed. 


###Should we ask Yue if it's ok to have multiple research questions that are seperate or do we need to tie them together with one central one?

