---
title: "Final Project:"
subtitle: "American Crime and Incarceration"
author: "PS I Love Yue: Oliver, Karam, Steven, Emely"
output: pdf_document
---

```{r packages-data, warning=FALSE, message=FALSE}
library(tidyverse)
crime <- read_csv("data/crime_and_incarceration_by_state.csv") 
```

```{r state_crime}

federal_crime <- crime %>%
  filter(jurisdiction == "FEDERAL")

state_crime <- crime %>%
  filter(jurisdiction != "FEDERAL")
```


Political rhetoric that has come to the forefront in recent decades is the idea 
of law and order, being tough on crime, and ensuring punishment to the fullest 
extent of the law. We went into this project knowing that the number of 
incarcerated individuals in the US is the highest of any other developed country 
with a similar economy and governmental system. Knowing this, we would assume 
that if the US has a strong belief that we need to be tough on crime, the high 
rates of incarceration are due to high crime rates, violent or otherwise. With 
this in mind, we wanted to find if the data would support this hypothesis.

```{r total-prisoners}
crime <- crime %>%
  group_by(year)%>%
  mutate(total_prisoner= sum(prisoner_count))
 
 crime %>%
  ggplot(crime, mapping= aes(x=year, y=total_prisoner))+
  geom_line()+
  labs(title = "Total US Prisoners",
  x = "Year", y = "Count") +
  theme_linedraw()
```
Above is the total prisoners in all 50 US states from 2001 to 2016. The graph 
shows a steady incline from 2001 to 2008 and then a steady decline followed by a 
sharp decline thereafter. Drawing attention to the range of the y-axis, we can 
see thar despite these sharp declines, the number of incarcerated individuals at 
the lowest point was still above 1.19 million, which is still infinitely higher 
than the rate of incarceration in other countries.

```{r total-violent-crime}
crime<- crime %>%
  filter(!is.na(violent_crime_total))%>%
  group_by(year)%>%
  mutate(total_violent_crime= sum(violent_crime_total))
 
 crime %>%
  ggplot(crime, mapping= aes(x=year, y=total_violent_crime))+
  geom_line()+
  labs(title = "Total Violent Crime in the US",
  x = "Year", y = "Count") +
  theme_linedraw()
```

```{r total-property-crime}
crime<- crime %>%
  filter(!is.na(property_crime_total))%>%
  group_by(year)%>%
  mutate(total_property_crime= sum(property_crime_total))
 
 crime %>%
  ggplot(crime, mapping= aes(x=year, y=total_property_crime))+
  geom_line()+
  labs(title = "Total Property Crime in the US", 
       subtitle = "Sum of Property Crime and Violent Crime", 
       x = "Year", 
       y = "Count") +
  theme_linedraw()
```


```{r total-crime}
crime<- crime %>%
  group_by(year)%>%
  mutate(total_crime= total_property_crime + total_violent_crime)
 
 crime %>%
  ggplot(crime, mapping= aes(x=year, y=total_crime))+
  geom_line()+
  labs(title = "Total Crime in the US",
  x = "Year", y = "Count") +
  theme_linedraw()
```

The three graphs above show the change in total violent crime 
(defined as murder/manslaughter, robbery, rape, and aggravated assault), 
total property crime (defined as burglary, larceny, and vehicle theft) and the 
total crime (sum of property and violent crime) committed between 2001 and 2016. 
What is immediately evident is that even when crime of all types was sharply 
decreasing, the number of incarcerated individuals was still steadily 
decreasing. Based on this alone, though we cannot definitively draw this 
conclusion, it would appear that rates of incarceration are actually not due to 
how much or how little crime is being committed and is instead due to other 
factors which we will explore later in our analyses. Possible explanations might 
be policy changes, new sentencing guidelines, amount of policing, and a plethora 
of other reasons.


```{r percent-pop}
crime<- crime %>%
  group_by(year)%>%
  mutate(percent_pop= sum(prisoner_count)/sum(state_population))
 
 crime %>%
  ggplot(crime, mapping= aes(x=year, y=percent_pop))+
  geom_line()+
  labs(title = "Percent Population Incarcerated",
  x = "Year", y = "Count") +
  theme_linedraw()
```
After drawing the previous conclusion, we thought there may be another 
explanation, even if crime was going down it could be that US population was 
increasing, when we graphed it above, it would appear that the percent of the 
US population which is imprisoned shows very strong similarities to the Total 
US Prison Population graph. However, that being said, even if the population is 
increasing, if the crime rate is staying the same, or decreasing as we observed
previously, there is still no valid reason to incarcerate more people.

### Steven

### Impact of Changing the Definition of Rape
```{r creation-of-rape-data-set}
crime_rape <- crime
crime_rape[is.na(crime_rape)] <- 0
```
 
```{r ggplot-of-rape-convictions}
crime_rape<- crime_rape %>%
  group_by(year)%>%
  mutate(total_rape_legacy= sum(rape_legacy))
 
crime_rape<- crime_rape %>%
  group_by(year)%>%
  mutate(total_rape_revised= sum(rape_revised))
 
  ggplot(data = crime_rape,
       mapping = aes(x = year)) +
  geom_line(aes(y = total_rape_legacy), color = "red") +
    geom_line(aes(y = total_rape_revised), color = "blue") +
    geom_vline(xintercept = 2012) +
    labs(title = "",
     x = "Year",
     y = "Number of Rape Cases",
     caption = "Red = Total Number of Prisoners convicted of rape
                      by the Legacy Definition in state custody
                Blue = Total Number of Prisoners convicted of rape
                        by the Revised Definition in state custody")
```
In 2012 the FBI changed their definition of rape from:
"The old definition was 'The carnal knowledge of a female forcibly and against
her will.' Many agencies interpreted this definition as excluding a long list
of sex offenses that are criminal in most jurisdictions, such as offenses
involving oral or anal penetration, penetration with objects, and rapes of
males.
 
to:
The new Summary definition of Rape is: 'Penetration, no matter how slight, of
the vagina or anus with any body part or object, or oral penetration by a sex
organ of another person, without the consent of the victim.'"
 
We wanted to see how the change in definition affected the number of
prisoners convicted for rape. Based on the graph, it seems like after the new
definition was implemented, the number of prisoners who are incarcerated for
rape increased. We will perform tests to find the p-value and try to
determine if this is a correlation relationship or a causal relationship.
 
https://ucr.fbi.gov/recent-program-updates/new-rape-definition-
frequently-asked-questions
 
### Proportions of Crime in the United States
 
```{r creating-proportion-variables}
crime_props <- crime %>%
  filter(jurisdiction != "FEDERAL") %>%
  mutate(violent_crime_total_prop = (violent_crime_total / prisoner_count)) %>%
  mutate(rape_legacy_prop = (rape_legacy / prisoner_count)) %>%
  mutate(murder_manslaughter_prop = (murder_manslaughter / prisoner_count)) %>%
  mutate(rape_revised_prop = (rape_revised / prisoner_count)) %>%
  mutate(robbery_prop = (robbery / prisoner_count)) %>%
  mutate(agg_asssault_prop = (agg_assault / prisoner_count)) %>%
  mutate(property_crime_total_prop = (property_crime_total / prisoner_count)) %>%
  mutate(burglary_prop = (burglary / prisoner_count)) %>%
  mutate(larceny_prop = (larceny / prisoner_count)) %>%
  mutate(vehicle_theft_prop = (vehicle_theft / prisoner_count))
 
crime_props <- crime_props %>%
  group_by(year) %>%
  mutate(total_rape_legacy_prop = (sum(rape_legacy_prop)/sum(prisoner_count))) %>%
  mutate(total_murder_manslaughter_prop = (sum(murder_manslaughter_prop)/sum(prisoner_count))) %>%
  mutate(total_rape_revised_prop = (sum(rape_revised_prop)/sum(prisoner_count))) %>%
  mutate(total_robbery_prop = (sum(robbery_prop)/sum(prisoner_count))) %>%
  mutate(total_agg_assault_prop = (sum(agg_asssault_prop)/sum(prisoner_count))) %>%
  mutate(total_burglary_prop = (sum(burglary_prop)/sum(prisoner_count))) %>%
  mutate(total_larceny_prop = (sum(larceny_prop)/sum(prisoner_count))) %>%
  mutate(total_vehicle_theft_prop = (sum(vehicle_theft_prop)/sum(prisoner_count))) %>%
  mutate(total_violent_crime_total = (sum(violent_crime_total_prop)/sum(prisoner_count))) %>%
  mutate(total_property_crime_total = (sum(property_crime_total_prop)/sum(prisoner_count)))
 
```
 
```{r proportion-of-crimes-including-larcency}
ggplot(data = crime_props, mapping = aes(x = year)) +
  geom_line(aes(y = total_rape_legacy_prop), color = "red",linetype="twodash") +
  geom_line(aes(y = total_murder_manslaughter_prop), color = "blue",linetype="twodash") +
  geom_line(aes(y = total_rape_revised_prop), color = "green",linetype="twodash") +
 geom_line(aes(y = total_robbery_prop), color = "yellow",linetype="twodash") +
  geom_line(aes(y = total_agg_assault_prop), color = "purple",linetype="twodash") +
  geom_line(aes(y = total_burglary_prop), color = "pink") +
  geom_line(aes(y = total_larceny_prop), color = "orange") +
  geom_line(aes(y = total_vehicle_theft_prop), color = "magenta")
  labs(title = "",
     x = "Year",
     y = "Proportion of prisoners convicted of each crime
            (not including federal incarcerated people)",
     caption = "Red = Proportion of Rape by legacy definiton
                Blue = Proportion of murder/manslaughter
                Green = Proportion of Rape by revised definition
                Yellow = Proportion of robbery
                Purple = Proportion of aggravated assault
                Pink = Proportion of burglary
                Orange = Proportion of larceny
                Magenta = Proportion of vehicle theft")
```
```{r proportion-of-crimes-including-larcency}
ggplot(data = crime_props, mapping = aes(x = year)) +
  geom_line(aes(y = total_rape_legacy_prop), color = "red", linetype="twodash") +
  geom_line(aes(y = total_murder_manslaughter_prop), color = "blue",linetype="twodash") +
  geom_line(aes(y = total_rape_revised_prop), color = "green",linetype="twodash") +
  geom_line(aes(y = total_robbery_prop), color = "yellow",linetype="twodash") +
  geom_line(aes(y = total_agg_assault_prop), color = "purple",linetype="twodash") +
  geom_line(aes(y = total_burglary_prop), color = "pink") +
  geom_line(aes(y = total_vehicle_theft_prop), color = "magenta") +
  labs(title = "",
     x = "Year",
     y = "Proportion of prisoners convicted of each crime
            (not including federal incarcerated people)",
     caption = "Red = Proportion of Rape by legacy definiton
                Blue = Proportion of murder/manslaughter
                Green = Proportion of Rape by revised definition
                Yellow = Proportion of robbery
                Purple = Proportion of aggravated assault
                Pink = Proportion of burglary
                Magenta = Proportion of vehicle theft")
```
 
```{r plot-violent-nonviolent-crimes}
ggplot(data = crime_props, mapping = aes(x = year)) +
  geom_line(aes(y = total_violent_crime_total), color = "red") +
  geom_line(aes(y = total_property_crime_total), color = "blue") +
  labs(title = "",
     x = "Year",
     y = "Proportion of prisoners convicted of each type of crime
            (not including federal incarcerated people)")
```
We wanted to see how the proportions of crimes were divided among the
prion population. Also we wanted to see whether violent or non-violent crimes
were represented more among the prison population. It appears that non violent
crimes make up a larger proportion of the prison population than violent crimes.
Generally, violent and non-violent crimes are decreasing the proportion of the
prison population. Nonviolent crimes seem to be decreasing at a faster rate
than violent crimes. Larceny makes up the largest proportion of crimes
in the prison population. So we took it out for the second visualization
to get a better look at the other crimes. Murder and manslaughter make up the
smallest proportion.

### Crime by year


```{r new-dataset}
totals_by_year <- state_crime %>%
  group_by(year) %>%
  summarise(violent_crime_total = sum(violent_crime_total),
            robbery = sum(robbery),
            agg_assault = sum(agg_assault),
            property_crime_total = sum(property_crime_total),
            burglary = sum(burglary),
            larceny = sum(larceny),
            vehicle_theft = sum(vehicle_theft))

# for some reason 2015 didn't get added to my dataset
```


```{r line-graph}


ggplot(data = totals_by_year,
       aes(x = year)) +
  geom_line(aes(y = violent_crime_total), color = "red") +
  geom_line(aes(y = robbery), color = "blue") +
  geom_line(aes(y = agg_assault), color = "green") +
  geom_line(aes(y = property_crime_total), color = "yellow") +
  geom_line(aes(y = burglary), color = "pink") +
  geom_line(aes(y = larceny), color = "black") +
  geom_line(aes(y = vehicle_theft), color = "purple")

#not a very effective visualization
  
```


```{r graphs, warning = FALSE, message = FALSE}

ggplot(data = totals_by_year, aes(x = year, y = violent_crime_total)) +
  geom_line() +
  geom_smooth(method = lm)

ggplot(data = totals_by_year, aes(x = year, y = robbery)) +
  geom_line() +
  geom_smooth(method = lm)

ggplot(data = totals_by_year, aes(x = year, y = agg_assault)) +
  geom_line() +
  geom_smooth(method = lm)

ggplot(data = totals_by_year, aes(x = year, y = property_crime_total)) +
  geom_line() +
  geom_smooth(method = lm)

ggplot(data = totals_by_year, aes(x = year, y = burglary)) +
  geom_line() +
  geom_smooth(method = lm)

ggplot(data = totals_by_year, aes(x = year, y = larceny)) +
  geom_line() +
  geom_smooth(method = lm)

ggplot(data = totals_by_year, aes(x = year, y = vehicle_theft)) +
  geom_line() +
  geom_smooth(method = lm)

```



```{r fitted-model}

violent_crime_total_model <- lm(violent_crime_total ~ year, 
                                data = totals_by_year)
tidy(violent_crime_total_model)


robbery_model <- lm(robbery ~ year, data = totals_by_year)
tidy(robbery_model)


agg_assault_model <- lm(agg_assault ~ year, data = totals_by_year)
tidy(agg_assault_model)


property_crime_total_model <- lm(property_crime_total ~ year, 
                                 data = totals_by_year)
tidy(property_crime_total_model)


burglary_model <- lm(burglary ~ year, data = totals_by_year)
tidy(burglary_model)


larceny_model <- lm(larceny ~ year, data = totals_by_year)
tidy(larceny_model)


vehicle_theft_model <- lm(vehicle_theft ~ year, data = totals_by_year)
tidy(vehicle_theft_model)

```


Here we attempted to answer whether different types of crime change together
over the years or if different crimes fluctuate in different years. We used 
data from the variables `violent_crime_total`, `robbery`, `agg_assault`,
`property_crime_total`, `burglary`, `larceny`, and `vehicle_theft`.

After attempting to compare the shapes of the crime vs. year graphs all together,
we found that the incredibly different scales make this task impossible to do
on a single graph. We then graphed these different crimes on different graphs
with linear models, making it easier to compare shape but still difficult to 
make objective claims about this data. One notable feature we noticed was that
there was a peak for many crimes between 2004 and 2008. We plan to research 
further to find possible causations for these trends.

Last, we created linear models for each crime to see how they each changed
between 2001 and 2016. Notice, all of these 7 crimes have a negative slope
meaning they overall decreased during this time period. The slopes vary in 
magnitude from about -7000 crimes/year to about -180,000 crimes/year.


