---
title: "Final Project:"
subtitle: "American Crime and Incarceration"
author: "PS I Love Yue: Oliver, Karam, Steven, Emely"
output: pdf_document
---

```{r packages-data, warning=FALSE, message=FALSE}
library(tidyverse)
crime <- read_csv("data/crime_and_incarceration_by_state.csv") 
view(crime)
```


### Beginning of Project

```{r ex-1}
#Seperate data between Federal prisoners and State prisoners
#Federal observations missing a lot of data so maybe best to work with state

federal_crime <- crime %>%
  filter(jurisdiction == "FEDERAL")

state_crime <- crime %>%
  filter(jurisdiction != "FEDERAL")
```


```{r ex-2}
#population and prisoner proportions

state_crime <- state_crime %>%
  mutate(prisoner_per_pop = prisoner_count/state_population)

state_crime %>%
  arrange(desc(prisoner_per_pop)) %>%
  select(jurisdiction, year, prisoner_count, state_population, prisoner_per_pop)

state_crime %>%
  arrange(state_population) %>%
  select(jurisdiction, year, state_population)

```
We can see here that the 25 highest prisoner per population observations 
are either Alaska or Delaware (two low population states, but not the lowest 
two).

Texas and Oklahoma are the only other states in the top 50 highest prisoner per 
population observations.


```{r ex-3}
#We can use this new data set to look at the most recent data

recent_state_crime <- state_crime %>%
  filter(year == 2016)

recent_state_crime %>%
  arrange(desc(prisoner_per_pop)) %>%
  select(jurisdiction, prisoner_per_pop)

recent_state_crime %>%
  arrange(desc(prisoner_count)) %>%
  select(jurisdiction, prisoner_count)


```
We can see here that in 2016 Oklahoma had the highest proportion of population 
imprisoned, while Texas had the highest raw number.

We also observe that the only states in the top 10 `prisoners_per_pop` and the 
top 10 `prisoner_count` are Texas, Florida, Arizona.

```{r}
recent_state_crime %>%
  ggplot(data = recent_state_crime, 
       mapping = aes(x = year,
                     y = prisoner_per_pop,
                     color = jurisdiction)) +
  geom_point() + 
  geom_smooth()
```




```{r ex-4}
crime<- crime %>%
  group_by(year)%>%
  mutate(total_murder= sum(murder_manslaughter))

crime<- crime %>%
  group_by(year)%>%
  mutate(total_vehicle_theft= sum(vehicle_theft))
  
 crime %>% 
  ggplot(crime, mapping= aes(x=year, y=total_murder))+
  geom_point()+
  labs(title = "title",
  x = "Year", y = "y") +
  theme_linedraw()
```
prison population

"violent_crime_total" is the sum of murder_manslaughter, rape_legacy, robbery, & agg_assault by year and state
"property_crime_total" is the sum of burglary, larceny, & vehicle_theft


possible question: correlation between population and prison population with bootstrapping/hypothesis test, comparing geographic region's prison populations 

```{r total-prisoners}
crime<- crime %>%
  group_by(year)%>%
  mutate(total_prisoner= sum(prisoner_count))
  
 crime %>% 
  ggplot(crime, mapping= aes(x=year, y=total_prisoner))+
  geom_line()+
  labs(title = "title",
  x = "Year", y = "Total US Prisoners") +
  theme_linedraw()
```
```{r total-property-crime}
crime<- crime %>%
  filter(!is.na(property_crime_total))%>%
  group_by(year)%>%
  mutate(total_crime= sum(property_crime_total))
  
 crime %>% 
  ggplot(crime, mapping= aes(x=year, y=total_crime))+
  geom_line()+
  labs(title = "title",
  x = "Year", y = "Total Property Crime in the US") +
  theme_linedraw()
```

```{r total-violent-crime}
crime<- crime %>%
  filter(!is.na(violent_crime_total))%>%
  group_by(year)%>%
  mutate(total_crime= sum(violent_crime_total))
  
 crime %>% 
  ggplot(crime, mapping= aes(x=year, y=total_crime))+
  geom_line()+
  labs(title = "title",
  x = "Year", y = "Total Violent Crime in the US") +
  theme_linedraw()
```

```{r total-violent-crime-by-state}
crime<- crime %>%
  filter(!is.na(violent_crime_total))%>%
  group_by(jurisdiction,year)%>%
  mutate(total_crime= sum(violent_crime_total))
  
 crime %>% 
  ggplot(crime, mapping= aes(x=year, y=total_crime))+
  geom_line()+
   facet_wrap(~jurisdiction)+
  labs(title = "title",
  x = "Year", y = "Total Violent Crime in the US") +
  theme_linedraw()
```

```{r population}

##not sure whats happening with this one
crime<- crime %>%
  filter(!is.na(state_population))%>%
  group_by(year)%>%
  mutate(pop= sum(state_population))

 crime %>% 
  ggplot(crime, mapping= aes(x=year, y=pop))+
  geom_line()+
  labs(title = "title",
  x = "Year", y = "Population") +
  theme_linedraw()
```



```{r percent-pop}
crime<- crime %>%
  group_by(year)%>%
  mutate(percent_pop= sum(prisoner_count)/sum(state_population))
  
 crime %>% 
  ggplot(crime, mapping= aes(x=year, y=percent_pop))+
  geom_line()+
  labs(title = "title",
  x = "Year", y = "Percent Population Incarcerated") +
  theme_linedraw()
```


```{r Prisoners%-by-state-2001}
crime<- crime %>%
  filter(year == 2001)%>%
  group_by(jurisdiction)%>%
  mutate(prisoner_2001= prisoner_count/state_population)
  
 crime %>% 
  ggplot(crime, mapping= aes(x= jurisdiction, y=prisoner_2001))+
  geom_col()+
  labs(title = "title",
  x = "Year", y = "prisoners 2001") +
  theme_linedraw()+ 
   theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

```{r}
crime <- crime %>% 
  mutate(region = case_when( 
  jurisdiction == "MAINE" ~ "Northeast",
jurisdiction == "NEW HAMPSHIRE" ~ "Northeast",
jurisdiction == "VERMONT" ~ "Northeast",
jurisdiction == "NEW YORK" ~ "Northeast",
jurisdiction == "CONNECTICUT" ~ "Northeast",
jurisdiction == "MASSACHUSETTS" ~ "Northeast",
jurisdiction == "RHODE ISLAND" ~ "Northeast",
jurisdiction == "PENNSYLVANIA" ~ "Northeast",
jurisdiction == "NORTH DAKOTA" ~ "North Central",
jurisdiction == "SOUTH DAKOTA" ~ "North Central",
jurisdiction == "NEBRASKA" ~ "North Central",
jurisdiction == "KANSAS" ~ "North Central",
jurisdiction == "MISSOURI" ~ "North Central",
jurisdiction == "IOWA" ~ "North Central",
jurisdiction == "MINNESOTA" ~ "North Central",
jurisdiction == "WISCONSIN" ~ "North Central",
jurisdiction == "ILLINOIS" ~ "North Central",
jurisdiction == "INDIANA" ~ "North Central",
jurisdiction == "MICHIGAN" ~ "North Central",
jurisdiction == "OHIO" ~ "North Central",
jurisdiction == "TEXAS" ~ "South",
jurisdiction == "OKLAHOMA" ~ "South",
jurisdiction == "LOUISIANA" ~ "South",
jurisdiction == "ARKANSAS" ~ "South",
jurisdiction == "MISSISSIPPI" ~ "South",
jurisdiction == "ALABAMA" ~ "South",
jurisdiction == "GEORGIA" ~ "South",
jurisdiction == "FLORIDA" ~ "South",
jurisdiction == "TENNESSEE" ~ "South",
jurisdiction == "SOUTH CAROLINA" ~ "South",
jurisdiction == "NORTH CAROLINA" ~ "South",
jurisdiction == "KENTUCKY" ~ "South",
jurisdiction == "VIRGINIA" ~ "South",
jurisdiction == "WEST VIRGINIA" ~ "South",
jurisdiction == "MARYLAND" ~ "South",
jurisdiction == "DELAWARE" ~ "South",
jurisdiction == "ALASKA" ~ "West",
jurisdiction == "HAWAII" ~ "West",
jurisdiction == "NEW MEXICO" ~ "West",
jurisdiction == "COLORADO" ~ "West",
jurisdiction == "WYOMING" ~ "West",
jurisdiction == "MONTANA" ~ "West",
jurisdiction == "IDAHO" ~ "West",
jurisdiction == "UTAH" ~ "West",
jurisdiction == "ARIZONA" ~ "West",
jurisdiction == "CALIFORNIA" ~ "West",
jurisdiction == "NEVADA" ~ "West",
jurisdiction == "OREGON" ~ "West",
jurisdiction == "WASHINGTON" ~ "West"))

 crime %>% 
  ggplot(crime, mapping= aes(x=year, y=percent_pop))+
  geom_line()+
  labs(title = "title",
  x = "Year", y = "Percent Population Incarcerated") +
  theme_linedraw()
```

